Part 1: Loading range size, abundance and morphological data and merging
```{r}
library(tidyverse)
library(openxlsx)
library(brms)
library(ape)

# Load range size data
range_size <- read.csv("../data/Bulbul_rangesize.csv")
str(range_size)

#Only keep columns 2-4 and columns 27-37
range_size <- range_size[, c(2:4, 27:37)]
str(range_size)

# Load abundance data
abundance <- read.csv("../data/abundances.csv")
str(abundance)

# Load morphological data if needed
morphology <- read.csv("../data/Bulbul_AVONET_data.csv")

# Merge datasets
merged_data <- merge(range_size, abundance, by.x = "Common.name", by.y = "Common.name") %>%
    distinct(Common.name, .keep_all = TRUE)
str(merged_data)
write.csv(merged_data, "../data/merged_range_abundance_data.csv")

# Load AVONET data
bulbul_specimens <- read.csv("../data/Bulbul_AVONET_data.csv", stringsAsFactors = FALSE)

#Check AviList column for unique species names
setdiff(
  unique(bulbul_specimens$AviList),
  unique(range_size$AviList)
)

setdiff(
  unique(range_size$AviList),
  unique(bulbul_specimens$AviList)
)

#Merge Scientific.name column of merged_data and the Species2_eBird column of bulbul_specimens
final_data <- merge(range_size, bulbul_specimens, by.x = "AviList", by.y = "AviList")
str(final_data)
bulbuls <- final_data

final_data <- merge(abundance, bulbul_specimens, by.x = "Scientific.name", by.y = "AviList")
str(final_data)
bulbuls <- final_data

# Save the final bulbul data
write.csv(bulbuls, "../data/Bulbul_final_data.csv")
```

Part 2: Calculate trait variances for species with 10+ bulbul_specimens
```{r}
bulbuls_count <- table(bulbuls$AviList)
bulbuls <- bulbuls[bulbuls$AviList %in% names(bulbuls_count[bulbuls_count >= 10]), ]
str(bulbuls)

# Calculate trait variance for each species (one row per species)
trait_cols <- c(
  "Beak.Length_Culmen", "Beak.Length_Nares",
  "Beak.Width", "Beak.Depth", "Tarsus.Length",
  "Wing.Length", "Kipps.Distance", "Secondary1",
  "Hand.wing.Index", "Tail.Length"
)

# Calculate variance for each trait, skipping NA values, for range size, habitat, trophic level, and lifestyle categories
bulbuls_var <- bulbuls %>%
  group_by(AviList, Range.Size) %>%
  summarise(
    across(all_of(trait_cols[1:length(trait_cols)]), 
           ~var(., na.rm = TRUE), 
           .names = "{.col}"),
    .groups = "drop"
  )
str(bulbuls_var)

#Calculate variance for abundance only
bulbuls_var <- bulbuls %>%
  group_by(Scientific.name, Abundance.estimate) %>%
  summarise(
    across(all_of(trait_cols[1:length(trait_cols)]), 
           ~var(., na.rm = TRUE), 
           .names = "{.col}"),
    .groups = "drop"
  )

write.csv(bulbuls_count, "../data/Bulbul_specimen_counts.csv")
```

Part 3: Plotting trait variances against range size and abundance
```{r}
library(ggplot2)

# Remove highest range size outlier for better visualization
#bulbuls_var <- bulbuls_var[bulbuls_var$Range.Size < 7000000, ]

# Plot trait variances against range size
ggplot(bulbuls_var, aes(x = log(Range.Size), y = Wing.Length)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Range size in square miles (log scale)", y = "σ² Wing Length") +
  theme_minimal()

ggsave("../results/WingLengthVar_vs_RangeSize_logged.png", width = 8, height = 5)

# Plot trait variances against abundance
ggplot(bulbuls_var, aes(x = log(Abundance.estimate), y = Tarsus.Length)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Abundance estimate (log scale)", y = "σ² Tarsus Length") +
  theme_minimal()
str(bulbuls_var)
```

Part 4: Run an LM to test relationship between trait variance and range size
```{r}
lm_wing_length <- lm(Tail.Length ~ Abundance.estimate, data = bulbuls_var)
summary(lm_wing_length)

boxplot(bulbuls_var$Wing.Length ~ bulbuls_var$Primary.Lifestyle, xlab = "Primary Lifestyle", ylab = "σ² Wing Length")

```

Part 5: Instead of variance, model morphospace area vs range size and abundance
```{r}
# Select morphological traits for PCA
pca_traits <- bulbuls %>%
  select(
    Beak.Length_Culmen,
    Tarsus.Length,
    Wing.Length,
    Tail.Length
  )

# Keep complete cases only
pca_complete <- pca_traits %>% drop_na()

# Keep matching metadata
meta_complete <- bulbuls %>%
  select(AviList, Range.Size) %>%
  slice(as.numeric(rownames(pca_complete)))

# Run PCA
pca_global <- prcomp(pca_complete, center = TRUE, scale. = TRUE)
summary(pca_global)

# Calculate morphospace area (convex hull area) for each species
pca_scores <- data.frame(
  AviList = meta_complete$AviList,
  Range.Size = meta_complete$Range.Size,
  PC1 = pca_global$x[, 1],
  PC2 = pca_global$x[, 2]
)

library(geometry)

morphospace_area <- pca_scores %>%
  group_by(AviList) %>%
  filter(n() >= 5) %>%   # enforce minimum sample size again
  summarise(
    morphospace_area = convhulln(
      cbind(PC1, PC2),
      options = "FA"
    )$vol,
    Range.Size    = mean(Range.Size, na.rm = TRUE),
    n_specimens          = n(),
    .groups = "drop"
  )

str(morphospace_area)
write.csv(morphospace_area, "../data/Bulbul_morphospace_areas.csv")

# Plot morphospace area vs range size
ggplot(morphospace_area,
       aes(x = log(Range.Size),
           y = morphospace_area)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal() +
  labs(
    x = "Range size: square miles (log scale)",
    y = "Morphospace area: PC1–PC2",
    title = "Species morphospace size vs geographic range size"
  )
```

Part 6: Time for brms modeling! But first, phylogeny building
```{r}
library(ape)

phylogeny <- read.tree("../data/Phylogeny/mcc_dated_clements_McTavish_2025.nex")
str(phylogeny)

phylogeny <- phylogeny[[6]]

tarsus_var$species_phylo <- gsub(" ", "_", tarsus_var$AviList)

species_keep <- unique(bulbuls$AviList_tree)

phylo_pruned <- drop.tip(
  phylogeny,
  setdiff(phylogeny$tip.label, species_keep)
)
str(phylo_pruned$edge.length)

# Save the pruned tree for later use
write.tree(phylo_pruned, "../data/Phylogeny/bulbul_phylogeny_pruned.nex")

#Load it again if needed
phylo_pruned <- read.tree("../data/Phylogeny/bulbul_phylogeny_pruned.nex")
str(phylo_pruned)

#Making covariance matrix for brms
A <- ape::vcv(phylo_pruned, corr = TRUE)

head(rownames(A))

#Now trying the brms model with the covariance matrix

bulbuls_var$Wing.Length_z <- scale(bulbuls_var$Wing.Length)
bulbuls_var$Tarsus.Length_z <- scale(bulbuls_var$Tarsus.Length)
bulbuls_var$Beak.Length_Culmen_z <- scale(bulbuls_var$Beak.Length_Culmen)
bulbuls_var$Tail.Length_z <- scale(bulbuls_var$Tail.Length)
bulbuls_var$Range.Size_z <- scale(log(bulbuls_var$Range.Size))

brm(
  Range.Size_z ~ Tail.Length_z +
    (1 | gr(species_phylo, cov = A)),
  data = bulbuls_var,
  data2 = list(A = A),
  chains = 4,
  cores = 4,
  iter = 40000
)
```

Part 7: Try a different method, do the random effect first and get residuals, then run the brms model with the residuals as the response variable
```{r}
#First, run the model with just the random effect to get the residuals
m <- lm(
  Beak.Length_Culmen ~ Sex + Measurer,
  data = bulbuls,
  na.action = na.exclude
)

bulbuls$Beak_resid <- residuals(m)

#Next, get variance for each species
beak_var <- bulbuls |>
  group_by(AviList) |>
  summarise(
    Beak.var = var(Beak_resid, na.rm = TRUE),
    n = n()
  )

#Next, merge with range size data
beak_var <- merge(beak_var, range_size, by.x = "AviList", by.y = "AviList")
str(beak_var)

beak_var$Range.Size_z <- scale(log(beak_var$Range.Size))
beak_var$Beak.var_z <- scale(beak_var$Beak.var)
names(beak_var)

beak_var$species_phylo <- gsub(" ", "_", beak_var$AviList)

#Finally, run the brms model with the residual variance as the response variable
brm(
  Range.Size_z ~ Beak.var_z + log(n) +
    (1 | gr(species_phylo, cov = A)),
  data = beak_var,
  data2 = list(A = A)
)

setdiff(beak_var$species_phylo, rownames(A))
setdiff(rownames(A), beak_var$species_phylo)
```
