```
Part 1: Load AVONET data and filter correctly for bulbul species
```{r}
AVONET <- read.csv("../data/AVONET raw data.csv", stringsAsFactors = FALSE)
str(AVONET)

bulbul_genera <- c("Pycnonotus", "Brachypodius", "Hypsipetes", "Iole",
       "Criniger", "Hemixos", "Alophoixus", "Rubigula",
       "Eurillas", "Setornis", "Tricholestes", "Phyllastrephus",
       "Bleda", "Andropadus", "Stelgidillas", "Calyptocichla", 
       "Neolestes", "Atimastillas", "Alcurus", "Nok", "Spizixos",
       "Ixonotus", "Thescelocichla", "Chlorocichla", "Baeopogon", 
       "Arizelocichla", "Acritillas", "Ixos", "Microtarsus")

AVONET$Species2_eBird <- trimws(as.character(AVONET$Species2_eBird))
pattern <- paste0("^(", paste(bulbul_genera, collapse = "|"), ")(\\b|\\s)")

idx <- !is.na(AVONET$Species2_eBird) & grepl(pattern, AVONET$Species2_eBird, ignore.case = TRUE)
bulbuls <- AVONET[idx, ]
str(bulbuls)
sum(unique(as.factor(bulbuls$Avibase.ID)) %in% unique(as.factor(AVONET$Avibase.ID)))

write.csv(bulbuls, "../data/Bulbul_AVONET_data.csv")
```

Part 2: Merge with BNHS data
```{r}
library(openxlsx)
BNHS <- read.xlsx("../data/BNHS_data.xlsx")
str(BNHS)

library(tidyverse)
BNHS$Specimen.number <- as.character(BNHS$Specimen.number)
BNHS$Publication <- as.character(BNHS$Publication)
bulbul_traits <- bind_rows(bulbuls, BNHS)
str(bulbul_traits)

write.csv(bulbul_traits, "../data/Bulbul_AVONET_data.csv")
```

Part 3: Load bulbul specimen data and model morphospace 
```{r}
bulbul_specimens <- read.csv("../data/Bulbul_AVONET_data.csv", stringsAsFactors = FALSE)
str(bulbul_specimens)

# Select traits to model morphospace
traits <- bulbul_specimens %>%
  select(
    Beak.Length_Culmen,
    Beak.Length_Nares,
    Beak.Width,
    Beak.Depth,
    Tarsus.Length,
    Wing.Length,
    Kipps.Distance,
    Tail.Length
  )

traits_complete <- traits %>% drop_na()

#Create metadata table
meta_complete <- bulbul_traits %>%
  select(Species2_eBird) %>%
  slice(as.numeric(rownames(traits_complete)))

# Perform PCA
pca_result <- prcomp(traits_complete, center = TRUE, scale. = TRUE)
summary(pca_result)

pca_df <- data.frame(
  PC1 = pca_result$x[, 1],
  PC2 = pca_result$x[, 2],
  Species2_eBird = meta_complete$Species2_eBird
)

# Plot PCA results
ggplot(pca_df, aes(PC1, PC2, colour = Species2_eBird)) +
  geom_point(alpha = 0.6, size = 1.8) +
  theme_classic() +
  labs(
    x = paste0("PC1 (", round(100 * summary(pca_result)$importance[2,1], 1), "%)"),
    y = paste0("PC2 (", round(100 * summary(pca_result)$importance[2,2], 1), "%)"),
    colour = "Species"
  ) +
  theme(
    legend.position = "none"  # IMPORTANT if many species
  )

ggsave("../results/Bulbul_Morphospace_PCA.png", width = 6, height = 5)

#Species centroids: OPTIONAL
species_centroids <- pca_df %>%
  group_by(Species2_eBird) %>%
  summarise(
    PC1 = mean(PC1),
    PC2 = mean(PC2),
    .groups = "drop"
  )

ggplot(pca_df, aes(PC1, PC2, colour = Species2_eBird)) +
  geom_point(alpha = 0.3, size = 1.5) +
  geom_point(
    data = species_centroids,
    aes(PC1, PC2),
    colour = "black",
    size = 3,
    inherit.aes = FALSE
  ) +
  theme_classic() +
  theme(legend.position = "none")
```

Part 4: One PCA for a species with wide range vs one with narrow range
```{r}
wide_range_species <- "Pycnonotus cafer"
narrow_range_species <- "Arizelocichla nigriceps"

wide_range_data <- bulbul_specimens %>%
  filter(Species2_eBird == wide_range_species) %>%
  select(
    Beak.Length_Culmen,
    Beak.Length_Nares,
    Beak.Width,
    Beak.Depth,
    Tarsus.Length,
    Wing.Length,
    Kipps.Distance,
    Tail.Length
  ) %>%
  drop_na()

narrow_range_data <- bulbul_specimens %>%
  filter(Species2_eBird == narrow_range_species) %>%
  select(
    Beak.Length_Culmen,
    Beak.Length_Nares,
    Beak.Width,
    Beak.Depth,
    Tarsus.Length,
    Wing.Length,
    Kipps.Distance,
    Tail.Length
  ) %>%
  drop_na()

combined_data <- rbind(wide_range_data, narrow_range_data)
pca_combined <- prcomp(combined_data, center = TRUE, scale. = TRUE)
summary(pca_combined)

pca_combined_df <- data.frame(
  PC1 = pca_combined$x[, 1],
  PC2 = pca_combined$x[, 2],
  Species2_eBird = c(rep(wide_range_species, nrow(wide_range_data)),
                     rep(narrow_range_species, nrow(narrow_range_data)))
)

hulls <- pca_combined_df %>%
  group_by(Species2_eBird) %>%
  slice(chull(PC1, PC2))

cb_palette <- c(
  "Pycnonotus cafer"     = "#E69F00",
  "Arizelocichla nigriceps" = "#009E73"
)

ggplot(pca_combined_df, aes(PC1, PC2, colour = Species2_eBird)) +
  geom_polygon(
    data = hulls,
    aes(fill = Species2_eBird),
    alpha = 0.25,
    colour = NA
  ) +
  geom_point(aes(shape = Species2_eBird), size = 2, alpha = 0.7) +
  scale_colour_manual(values = cb_palette) +
  scale_fill_manual(values = cb_palette) +
  theme_classic() +
  labs(
    x = paste0("PC1 (", round(100 * summary(pca_combined)$importance[2,1], 1), "%)"),
    y = paste0("PC2 (", round(100 * summary(pca_combined)$importance[2,2], 1), "%)"),
    colour = "Species",
    fill   = "Species",
    shape  = "Species"
  ) +
  theme(legend.position = "top")

ggsave("../results/Bulbul_Wide_vs_Narrow_Range_PCA.png", width = 6, height = 5)
```