Part 1: Load Bulbul_AVONET_data, AVONET_alldupes and Rob Barber data files
```{r}
library(tidyverse)
library(openxlsx)
library(stringr)

bulbuls <- read.csv("../data/Bulbul_AVONET_data.csv")
str(bulbuls)
alldupes <- read.xlsx("../data/Duplicates/AVONET_alldupes.xlsx")
str(alldupes)
rob <- read.xlsx("../data/Duplicates/Rob_Barber_bulbul_data.xlsx")
str(rob)

#Checking bulbuls object to find duplicate specimen numbers
duplicated_bulbuls <- bulbuls %>%
        group_by(Specimen.number) %>%
        summarise(count = n()) %>%
        filter(count > 1)

#Find all duplicated_bulbuls$Specimen.number that are NOT present in alldupes$
unique_duplicated_bulbuls <- duplicated_bulbuls %>%
        filter(!Specimen.number %in% alldupes$Specimen.number)

#Remove first and last row of unique_duplicated_bulbuls
unique_duplicated_bulbuls <- unique_duplicated_bulbuls[-c(1, nrow(unique_duplicated_bulbuls)), ]
```
Part 2: Make a separate list of rows called ER_dupes from the bulbuls dataset of specimen numbers present in unique_duplicated_bulbuls
```{r}
er_dupes <- bulbuls |>
        filter(Specimen.number %in% unique_duplicated_bulbuls$Specimen.number)
str(er_dupes)
```
Part 3: Now add all the Rob Barber data to the ER_dupes dataset
```{r}
rob <- rob |>
  rename(
    AviList = `Scientific_Name.(BirdTree)`,
    Specimen.number = Number,
    Beak.Length_Culmen = Bill_TotalCulmen,
    Beak.Length_Nares = Bill_FromNostril,
    Beak.Width = Bill_Width,
    Beak.Depth = Bill_Depth,
    Gape_Width = Gape.Width,
    Tarsus.Length = Tarsus_Length,
    Kipps.Distance = `Kipp's_Distance`,
    Wing.Length = Wing_Chord,
    Tail.Length = Tail_Length
  )
rob$Subspecies <- as.character(rob$Subspecies)
rob$Comments <- as.character(rob$Comments)

er_dupes <- bind_rows(er_dupes, rob) |>
  select(any_of(intersect(names(er_dupes), names(rob))))
str(er_dupes)
write.csv(er_dupes, "../data/Duplicates/ER_dupes.csv", row.names = FALSE)
```
Part 3: Now load ER_dupes again and group correctly by Specimen.number
```{r}
er_dupes <- read.csv("../data/Duplicates/ER_dupes.csv")
str(er_dupes)
er_dupes <- er_dupes |>
  group_by(Specimen.number, AviList) |>
  arrange(Specimen.number, AviList, .by_group = TRUE)
write.csv(er_dupes, "../data/Duplicates/ER_dupes.csv", row.names = FALSE)
```
Part 4: Bind rows with all_dupes and fill in the blanks
```{r}
er_dupes_clean <- er_dupes %>%
  mutate(
    # Create Unique.Specimen.Identifier
    Unique.Specimen.Identifier = paste(AviList, Specimen.number, sep = "_"),
    
    # Split genus + species if needed
    Species1 = word(AviList, 2),
    
    # Fill columns that exist in target but not here
    Order1 = NA,
    Sequence = NA,
    Source = "NHMUK",
    Secondary1 = Wing.Length - Kipps.Distance,
    `Hand-wing.Index` = NA,
    Sample.Number = NA,
    Avonet.Protocol.Used = NA,
    Protocol.both.pairs = NA,
    Protocol.neither.pair = NA
  ) %>%
  
  # Rename columns to match target
  rename(
    `Specimen.number` = Specimen.number,
    `Sex` = Sex,
    `Locality` = Locality,
    `Country` = Country,
    `Beak.Length_Culmen` = Beak.Length_Culmen,
    `Beak.Length_Nares` = Beak.Length_Nares,
    `Beak.Width` = Beak.Width,
    `Beak.Depth` = Beak.Depth,
    `Tarsus.Length` = Tarsus.Length,
    `Wing.Length` = Wing.Length,
    `Kipps.Distance` = Kipps.Distance,
    `Tail.Length` = Tail.Length,
    `Measurer` = Measurer
  )
str(er_dupes_clean)
er_dupes_clean <- as.data.frame(er_dupes_clean)
# Step 2: Ensure same column order as duplicates dataset
er_dupes_clean <- er_dupes_clean %>%
  select(names(alldupes))  # 'duplicates' is your target dataset

# Step 3: Bind
combined_data <- bind_rows(alldupes, er_dupes_clean)
str(combined_data)
write.csv(combined_data, "../data/Duplicates/combined_dupes.csv", row.names = FALSE)
```
Part 5: Now load combined_dupes and sort everything alphabetically by Order, then Family, then Species within that, after which we extract a bulbul_dupes file for all Pycnonotidae
```{r}
combined_data <- read.csv("../data/Duplicates/combined_dupes.csv")
str(combined_data)
combined_data <- combined_data %>%
    arrange(Order1, Family, Unique.Specimen.Identifier)
write.csv(combined_data, "../data/Duplicates/combined_dupes_sorted.csv", row.names = FALSE)

bulbul_dupes <- combined_data %>%
    filter(Family == "Pycnonotidae")
str(bulbul_dupes)
write.csv(bulbul_dupes, "../data/Duplicates/bulbul_dupes.csv", row.names = FALSE)
```
Next step: Correlations and repeatability of measurements for duplicated specimens
```{r}
bulbul_dupes <- read.csv("../data/Duplicates/Bulbul_Duplicates.csv")
str(bulbul_dupes)

library(dplyr)
library(tidyr)
library(rptR)
library(ggplot2)

# -----------------------------
# 1. Define traits
# -----------------------------
traits <- c(
  "Beak.Length_Culmen",
  "Beak.Length_Nares",
  "Beak.Width",
  "Beak.Depth",
  "Tarsus.Length",
  "Wing.Length",
  "Kipps.Distance",
  "Tail.Length"
)

# -----------------------------
# 2. Repeatability (rptR)
# -----------------------------
rpt_results <- lapply(traits, function(traits) {
  rpt(
    as.formula(paste(traits, "~ (1|Unique.Specimen.Identifier)")),
    grname = "Unique.Specimen.Identifier",
    data = bulbul_dupes,
    datatype = "Gaussian",
    nboot = 1000,
    npermut = 500
  )
})

names(rpt_results) <- traits

# Extract repeatability estimates and 95% confidence intervals
rpt_ci <- do.call(rbind, lapply(names(rpt_results), function(traits) {
    data.frame(
        Trait = traits,
        R = rpt_results[[traits]]$R,
        CI_lower = rpt_results[[traits]]$CI_emp$`2.5%`,
        CI_upper = rpt_results[[traits]]$CI_emp$`97.5%`
    )
}))
write.csv(rpt_ci, "../Results/Sensitivity/Bulbul_Duplicates_Repeatability.csv", row.names = FALSE)

# -----------------------------
# 3. Extract bootstrap values
# -----------------------------
boot_df <- do.call(rbind, lapply(names(rpt_results), function(traits) {
  
  boot_vals <- rpt_results[[traits]]$R_boot[,1]
  
  data.frame(
    Trait = traits,
    R = boot_vals
  )
}))

# -----------------------------
# 4. GLM across traits
# -----------------------------
# (tests whether repeatability differs among traits)

glm_model <- glm(R ~ Trait, data = boot_df)

summary(glm_model)
anova(glm_model, test = "F")

# -----------------------------
# 5. Reshape to wide format
# -----------------------------
bulbul_wide <- bulbul_dupes %>%
  arrange(Unique.Specimen.Identifier, Sample.Number) %>%
  pivot_wider(
    id_cols = Unique.Specimen.Identifier,
    names_from = Sample.Number,
    values_from = all_of(traits)
  )

# -----------------------------
# 6. Pearson correlation
# -----------------------------
results_df <- tibble(
    Trait = traits,
    Pearson_Correlation = sapply(traits, function(trait) {
        x <- bulbul_wide[[paste0(trait, "_1")]]
        y <- bulbul_wide[[paste0(trait, "_2")]]
        
        cor(x, y, use = "complete.obs")
    }),
    CV = sapply(traits, function(trait) {
        x <- bulbul_wide[[paste0(trait, "_1")]]
        y <- bulbul_wide[[paste0(trait, "_2")]]
        
        means <- rowMeans(cbind(x, y), na.rm = TRUE)
        sds <- apply(cbind(x, y), 1, sd, na.rm = TRUE)
        cv <- sds / means
        
        mean(cv, na.rm = TRUE)
    })
)

results_df
write.csv(results_df, "../Results/Sensitivity/Bulbul_Duplicates_Correlations.csv", row.names = FALSE)

# -----------------------------
# 8. Optional: protocol comparison
# -----------------------------
bulbul_dupes$ProtocolGroup <- ifelse(
  bulbul_dupes$Protocol.both.pairs == 1,
  "Both_AVONET",
  "Mixed_or_nonAVONET"
)

# Example: split and recompute correlations
cor_by_protocol <- bulbul_dupes %>%
  group_split(ProtocolGroup) %>%
  setNames(unique(bulbul_dupes$ProtocolGroup)) %>%
  lapply(function(df) {
    
    wide <- df %>%
      arrange(Unique.Specimen.Identifier, Sample.Number) %>%
      pivot_wider(
        id_cols = Unique.Specimen.Identifier,
        names_from = Sample.Number,
        values_from = all_of(traits)
      )
    
    sapply(traits, function(trait) {
      cor(
        wide[[paste0(trait, "_1")]],
        wide[[paste0(trait, "_2")]],
        use = "complete.obs"
      )
    })
  })

cor_by_protocol

# -----------------------------
# 9. Plot (example for one trait)
# -----------------------------
ggplot(bulbul_wide, aes(
  x = Tail.Length_1,
  y = Tail.Length_2
)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme_classic() +
  labs(
    x = "Measurement 1",
    y = "Measurement 2",
    title = "Repeatability of Tail Length"
  )

ggsave("../Results/Sensitivity/Bulbuls_Tail_Length_Repeatability.png", width = 6, height = 6)
```
Now let's just do the same for all oscines to keep in hand
```{r}
avonet_dupes <- read.csv("../data/Duplicates/AVONET_Duplicates.csv")
str(avonet_dupes)

suboscine_families <- c(
  # Acanthisitti
  "Acanthisittidae",
  
  # Old World suboscines
  "Philepittidae",
  "Eurylaimidae",
  "Calyptomenidae",
  "Sapayoidae",
  "Pittidae",
  
  # New World suboscines
  "Melanopareiidae",
  "Conopophagidae",
  "Thamnophilidae",
  "Grallariidae",
  "Rhinocryptidae",
  "Formicariidae",
  "Furnariidae",
  "Pipridae",
  "Cotingidae",
  "Tityridae",
  "Oxyruncidae",
  "Onychorhynchidae",
  "Tyrannidae"
)

oscine_dupes <- subset(
  avonet_dupes,
  Order1 == "PASSERIFORMES" &
  !Family %in% suboscine_families
)
str(oscine_dupes)
unique(oscine_dupes$Family)

write.csv(oscine_dupes, "../data/Duplicates/AVONET_oscine_duplicates.csv", row.names = FALSE)

#Now load it and do the process again
oscine_dupes <- read.csv("../data/Duplicates/AVONET_oscine_duplicates.csv")
str(oscine_dupes)

# -----------------------------
# 1. Define traits
# -----------------------------
traits <- c(
  "Beak.Length_Culmen",
  "Beak.Length_Nares",
  "Beak.Width",
  "Beak.Depth",
  "Tarsus.Length",
  "Wing.Length",
  "Kipps.Distance",
  "Tail.Length"
)

# -----------------------------
# 2. Repeatability (rptR)
# -----------------------------
rpt_results <- lapply(traits, function(trait) {
    rpt(
        as.formula(paste(trait, "~ (1|Unique.Specimen.Identifier) + (1|Family)")),
        grname = c("Unique.Specimen.Identifier", "Family"),
        data = oscine_dupes,
        datatype = "Gaussian",
        nboot = 1000,
        npermut = 500
    )
})

names(rpt_results) <- traits

# Extract repeatability estimates and 95% confidence intervals
rpt_ci <- do.call(rbind, lapply(names(rpt_results), function(trait) {
  data.frame(
    Trait = trait,
    R_Specimen = rpt_results[[trait]]$R["Unique.Specimen.Identifier"],
    R_Family = rpt_results[[trait]]$R["Family"],
    CI_lower = rpt_results[[trait]]$CI_emp$`2.5%`[1],
    CI_upper = rpt_results[[trait]]$CI_emp$`97.5%`[1]
  )
}))

write.csv(rpt_ci, "../Results/Sensitivity/oscine_Duplicates_Repeatability.csv", row.names = FALSE)

# -----------------------------
# 3. Extract bootstrap values
# -----------------------------
boot_df <- do.call(rbind, lapply(names(rpt_results), function(traits) {
  
  boot_vals <- rpt_results[[traits]]$R_boot[,1]
  
  data.frame(
    Trait = traits,
    R = boot_vals
  )
}))

# -----------------------------
# 4. GLM across traits
# -----------------------------
# (tests whether repeatability differs among traits)

glm_model <- glm(R ~ Trait, data = boot_df)

summary(glm_model)
anova(glm_model, test = "F")

# -----------------------------
# 5. Reshape to wide format
# -----------------------------
oscine_wide <- oscine_dupes %>%
  arrange(Unique.Specimen.Identifier, Sample.Number) %>%
  pivot_wider(
    id_cols = Unique.Specimen.Identifier,
    names_from = Sample.Number,
    values_from = all_of(traits)
  )

# -----------------------------
# 6. Pearson correlation
# -----------------------------
results_df <- tibble(
    Trait = traits,
    Pearson_Correlation = sapply(traits, function(trait) {
        x <- oscine_wide[[paste0(trait, "_1")]]
        y <- oscine_wide[[paste0(trait, "_2")]]
        
        cor(x, y, use = "complete.obs")
    }),
    CV = sapply(traits, function(trait) {
        x <- oscine_wide[[paste0(trait, "_1")]]
        y <- oscine_wide[[paste0(trait, "_2")]]
        
        means <- rowMeans(cbind(x, y), na.rm = TRUE)
        sds <- apply(cbind(x, y), 1, sd, na.rm = TRUE)
        cv <- sds / means
        
        mean(cv, na.rm = TRUE)
    })
)

results_df
write.csv(results_df, "../Results/Sensitivity/oscine_Duplicates_Correlations.csv", row.names = FALSE)

# -----------------------------
# 8. Optional: protocol comparison
# -----------------------------
oscine_dupes$ProtocolGroup <- ifelse(
  oscine_dupes$Protocol.both.pairs == 1,
  "Both_AVONET",
  "Mixed_or_nonAVONET"
)

# Example: split and recompute correlations
cor_by_protocol <- oscine_dupes %>%
  group_split(ProtocolGroup) %>%
  setNames(unique(oscine_dupes$ProtocolGroup)) %>%
  lapply(function(df) {
    
    wide <- df %>%
      arrange(Unique.Specimen.Identifier, Sample.Number) %>%
      pivot_wider(
        id_cols = Unique.Specimen.Identifier,
        names_from = Sample.Number,
        values_from = all_of(traits)
      )
    
    sapply(traits, function(trait) {
      cor(
        wide[[paste0(trait, "_1")]],
        wide[[paste0(trait, "_2")]],
        use = "complete.obs"
      )
    })
  })

cor_by_protocol

# -----------------------------
# 9. Plot (example for one trait)
# -----------------------------
ggplot(oscine_wide, aes(
  x = Tail.Length_1,
  y = Tail.Length_2
)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme_classic() +
  labs(
    x = "Measurement 1",
    y = "Measurement 2",
    title = "Repeatability of Tail Length"
  )

ggsave("../Results/Sensitivity/oscines_Tail_Length_Repeatability.png", width = 6, height = 6)
```
One more analysis: Checking how many specimens are required to capture 90% of the variation in each trait for the 10 most sampled species
```{r}
bulbuls <- read.csv("../data/Bulbul_AVONET_data_integrated.csv")
str(bulbuls)

traits <- c(
  "Beak.Length_Culmen",
  "Beak.Length_Nares",
  "Beak.Width",
  "Beak.Depth",
  "Tarsus.Length",
  "Wing.Length",
  "Kipps.Distance",
  "Tail.Length"
)

top_species <- bulbuls %>%
  group_by(AviList) %>%
  summarise(n = sum(!is.na(Beak.Length_Culmen))) %>%  # pick a well-sampled trait
  arrange(desc(n)) %>%
  slice(1:10) %>%
  pull(AviList)

cv_fun <- function(x) {
  x <- x[!is.na(x)]
  if(length(x) < 2) return(NA)
  sd(x) / mean(x)
}

full_cv <- bulbuls %>%
  filter(AviList %in% top_species) %>%
  group_by(AviList) %>%
  summarise(across(all_of(traits), cv_fun, .names = "CV_{.col}"))

#Permutation/subsampling function
sample_cv <- function(data, trait, n, n_iter = 100) {
  
  vals <- data[[trait]]
  vals <- vals[!is.na(vals)]
  
  if(length(vals) < n) return(NA)
  
  replicate(n_iter, {
    sample_vals <- sample(vals, n)
    sd(sample_vals) / mean(sample_vals)
  })
}

library(purrr)

results <- list()

for(sp in top_species) {
  
  sp_data <- bulbuls %>% filter(AviList == sp)
  
  for(trait in traits) {
    
    vals <- sp_data[[trait]]
    vals <- vals[!is.na(vals)]
    
    if(length(vals) < 10) next  # skip poorly sampled
    
    full_val <- cv_fun(vals)
    
    sizes <- seq(5, length(vals), by = 5)
    
    for(n in sizes) {
      
      cv_samples <- sample_cv(sp_data, trait, n, n_iter = 100)
      
      results[[length(results)+1]] <- data.frame(
        Species = sp,
        Trait = trait,
        SampleSize = n,
        CV_mean = mean(cv_samples, na.rm = TRUE),
        CV_sd = sd(cv_samples, na.rm = TRUE),
        Full_CV = full_val,
        Proportion = mean(cv_samples, na.rm = TRUE) / full_val
      )
    }
  }
}

results_df <- do.call(rbind, results)

#Can set alternate thresholds as well, default is 0.95 for 95% of the variation captured, but could also look at 0.90 or 0.99 for example
threshold <- 0.95

min_sample_sizes <- results_df %>%
  group_by(Species, Trait) %>%
  filter(Proportion >= threshold) %>%
  summarise(Min_N = min(SampleSize), .groups = "drop")

#Now to calculate mean minimum sample size per trait
mean_min_sample_size <- min_sample_sizes %>%
  group_by(Trait) %>%
  summarise(Mean_Min_N = mean(Min_N, na.rm = TRUE), .groups = "drop")

write.csv(min_sample_sizes, "../Results/Sensitivity/Bulbul_Sample_Size_CV.csv", row.names = FALSE)
write.csv(mean_min_sample_size, "../Results/Sensitivity/Bulbul_Mean_Traitwise_Sample_Size_CV.csv", row.names = FALSE)

ggplot(results_df, aes(x = SampleSize, y = Proportion)) +
  geom_line() +
  geom_hline(yintercept = 0.95, linetype = "dashed") +
  facet_grid(Trait ~ Species, scales = "free_y") +
  theme_bw()

ggsave("../Results/Sensitivity/Bulbul_Sample_Size_CV.png", width = 12, height = 8)
```