##Poisson models

Loading fisheries data:
```{r}
install.packages("MASS")
install.packages("ggpubr")
require(ggplot2)
require(MASS)
require(ggpubr)
fish <- read.csv("../data/fisheries.csv", stringsAsFactors = T)
str(fish)
```

Initial plotting:
```{r}
ggplot(fish, aes(x = MeanDepth, y = TotAbund)) +
    geom_point() +
    labs(x = "Mean Depth (km)", y = "Total abundance") +
    theme_classic()
```

GLM for total abundance explained by Mean Depth, and its summary
```{r}
M1 <- glm(formula = TotAbund ~ MeanDepth, family = "poisson", data = fish)
summary(M1)
```

Plotting the model diagnostics:
```{r}
par(mfrow = c(2,2))
plot(M1)
sum(cooks.distance(M1)>1)
```

Scatterplot with year as fixed factor just to check:
```{r}
scatterplot<-ggplot(fish, aes(x=MeanDepth, y=TotAbund, color=factor(Period)))+
  geom_point()+
  labs(x= "Mean Depth (km)", y="Total Abundance")+
  theme_classic()+
  scale_color_discrete(name="Period", labels=c("1979-1989", "1997-2002"))
boxplot<- ggplot(fish, aes(x=factor(Period, labels=c("1979-1989", "1997-2002")), y=TotAbund))+
  geom_boxplot()+
  theme_classic()+
  labs(x="Period", y="Total Abundance")
ggarrange(scatterplot, boxplot, labels=c("A","B"), ncol=1, nrow=2)
```

Adding period as a fixed factor:
```{r}
fish$Period <- factor(fish$Period)
M2 <- glm(TotAbund ~ MeanDepth * Period, data = fish, family = "poisson")
summary(M2)
```

Dispersion is still >100 which is too much, time for a negative binomial:
```{r}
M3 <- glm.nb(TotAbund ~ MeanDepth * Period, data = fish)
summary(M3)
```

The multiplicative interaction is not needed as the MeanDepth:Period interaction is not significant:
```{r}
M4 <- glm.nb(formula = TotAbund ~ MeanDepth + Period, data = fish)
summary(M4)
```

Now let's look at the model diagnostics for M4 and the dispersion paramater:
```{r}
par(mfrow = c(2,2))
plot(M4)
```

Now let's do a final plot with confidence intervals and standard error:
The reason for using predict is basically to get the model values for the distribution of period 1 and period 2 points to calculate confidence intervals
```{r}
period1 <- data.frame(MeanDepth=seq(from=0.804, to=4.865, length=100), Period="1")
period2 <- data.frame(MeanDepth=seq(from=0.804, to=4.865, length=100), Period="2")
period1_predictions<- predict(M4, newdata = period1, type = "link", se.fit = TRUE) # the type="link" here predicted the fit and se on the log-linear scale. 
period2_predictions<- predict(M4, newdata = period2, type = "link", se.fit = TRUE)
period1$pred<- period1_predictions$fit
period1$se<- period1_predictions$se.fit
period1$upperCI<- period1$pred+(period1$se*1.96)
period1$lowerCI<- period1$pred-(period1$se*1.96)
period2$pred<- period2_predictions$fit
period2$se<- period2_predictions$se.fit
period2$upperCI<- period2$pred+(period2$se*1.96)
period2$lowerCI<- period2$pred-(period2$se*1.96)
complete<- rbind(period1, period2)

# Making the Plot 
ggplot(complete, aes(x=MeanDepth, y=exp(pred)))+ 
  geom_line(aes(color=factor(Period)))+
  geom_ribbon(aes(ymin=exp(lowerCI), ymax=exp(upperCI), fill=factor(Period), alpha=0.3), show.legend = FALSE)+ 
  geom_point(fish, mapping = aes(x=MeanDepth, y=TotAbund, color=factor(Period)))+
  labs(y="Total Abundance", x="Mean Depth (km)")+
  theme_classic()+
  scale_color_discrete(name="Period", labels=c("1979-1989", "1997-2002"))

```

Just do all this much easier:
```{r}
require(ggeffects)
plot(ggpredict(M4, terms=c("MeanDepth", "Period")), show_data=T)
```

Moving on to the bee mites example:
Fitting the Poisson model after loading bee mites data
```{r}
mites <- read.csv("../data/bee_mites.csv")
mites_m1 <- glm(formula = Dead_mites ~ Concentration, data = mites, family = "poisson")
summary(mites_m1)
```

Exercise for Galapagos data gala.txt
```{r}
gala <- read.delim("../data/gala.txt")
summary(gala)
island_number <- glm(formula = Species ~ log(Area), data = gala, family = "poisson")
summary(island_number)
```

Instead of poisson, try quasi-poisson or negative binomial:
```{r}
island_number <- glm(formula = Species ~ log(Area), data = gala, family = "quasipoisson")
summary(island_number)

island_number <- glm.nb(formula = Species ~ log(Area), data = gala)
summary(island_number)
```
Negative binomial gives a dispersion parameter of around 1, which is acceptable
Now to check the residuals vs fitted
```{r}
island_number <- glm(formula = Species ~ log(Area), data = gala, family = "quasipoisson")
summary(island_number)

island_number <- glm.nb(formula = Species ~ log(Area), data = gala)
summary(island_number)

par(mfrow = c(2,2))
plot(island_number)
```

Question2: Amphibian roadkill
```{r}
amphib <- read.delim("../data/RoadKills.txt")
summary(amphib)
amphib$D.PARK <- amphib$D.PARK/1000
amphib_park <- glm(formula = TOT.N ~ D.PARK, data = amphib, family = "poisson")
summary(amphib_park)
```

Negative binomial and quasipoisson:
```{r}
amphib_park <- glm(formula = TOT.N ~ D.PARK, data = amphib, family = "quasipoisson")
summary(amphib_park)
amphib_park <- glm.nb(formula = TOT.N ~ D.PARK, data = amphib)
summary(amphib_park)
```


Now we move to Handout GLM H02
##Binomal and binary GLMs

```{r}
require(ggplot2)
require(ggpubr)
worker <- read.csv("../data/workerbees.csv", stringsAsFactors = T)
str(worker)

scatterplot <- ggplot(worker, aes(x = CellSize, y = Parasites))+
  geom_point()+
  labs(x = "Cell Size (cm)", y = "Probability of Parasite")+
  theme_classic()
boxplot <- ggplot(worker, aes(x = factor(Parasites), y = CellSize))+
  geom_boxplot()+
  theme_classic()+
  labs(x = "Presence/Absence of Parasites", y = "Cell Size (cm)")
ggarrange(scatterplot, boxplot, labels = c("A","B"), ncol = 1, nrow = 2)
```

Fitting the model to look at whether cell size affects parasites:
```{r}
M1 <- glm(Parasites ~ CellSize, data = worker, family = "binomial")
summary(M1)
```

Plotting the GLM you just made:
```{r}
range(worker$CellSize)
new_data <- data.frame(CellSize = seq(from = 0.352, to = 0.664, length = 100))
predictions <- predict(M1, newdata = new_data, type = "link", se.fit = TRUE)
new_data$pred<- predictions$fit
new_data$se<- predictions$se.fit
new_data$upperCI<- new_data$pred+(new_data$se*1.96)
new_data$lowerCI<- new_data$pred-(new_data$se*1.96)

ggplot(new_data, aes(x=CellSize, y=plogis(pred)))+ 
  geom_line(col="black")+
  geom_point(worker, mapping = aes(x=CellSize, y=Parasites), col="blue")+
  geom_ribbon(aes(ymin=plogis(lowerCI), ymax=plogis(upperCI), alpha=0.2), show.legend = FALSE)+ 
  labs(y="Probability of Infection", x="Cell Size (cm)")+
  theme_classic()
```

New example to investigate: Chytrid infection status in the Pyrenees:
```{r}
require(ggplot2)
chytrid<- read.csv("../data/chytrid.csv", stringsAsFactors = T)
str(chytrid)

scatterplot<-ggplot(chytrid, aes(x=Springavgtemp, y=InfectionStatus))+
  geom_point()+
  labs(x= "Probability of Infection", y="Average Spring Temperature (Degrees Celsius)")+
  theme_classic()
boxplot<- ggplot(chytrid, aes(x=factor(InfectionStatus), y=Springavgtemp))+
  geom_boxplot()+
  theme_classic()+
  labs(x="Presence/Absence of Infection", y="Average Spring Temperature (Degrees Celsius)")
ggarrange(scatterplot, boxplot, labels=c("A","B"), ncol=1, nrow=2)
```

Fitting the GLM onto the data to investigate whether probability of fungus infection increases with spring Temperature:
```{r}
M2 <- glm(InfectionStatus ~ Springavgtemp, data = chytrid, family = "binomial")
summary(M2)

#Plotting
new_data <- data.frame(Springavgtemp=seq(from=0.99, to=13.67, length=100))
predictions<- predict(M2, newdata = new_data, type = "link", se.fit = TRUE) # the type="link" here predicted the fit and se on the log-linear scale. 
new_data$pred<- predictions$fit
new_data$se<- predictions$se.fit
new_data$upperCI<- new_data$pred+(new_data$se*1.96)
new_data$lowerCI<- new_data$pred-(new_data$se*1.96)

# Making the Plot 
ggplot(new_data, aes(x=Springavgtemp, y=plogis(pred)))+ 
  geom_line(col="black")+
  geom_point(chytrid, mapping = aes(x=Springavgtemp, y=InfectionStatus), col="blue")+
  geom_ribbon(aes(ymin=plogis(lowerCI), ymax=plogis(upperCI), alpha=0.2), show.legend = FALSE)+ 
  labs(y="Probability of Infection", x="Average Spring Temperature (Degrees Celsius)")+
  theme_classic()
```

The model just isn't very good fitting and the psuedo-R squared is very bad, so we convert to binomial outcomes
Count the number of positives, number of negatives to make it into proportions:
```{r}
chytrid_binomial<- read.csv("../data/chytrid_binomial.csv", stringsAsFactors = T)
str(chytrid_binomial)

M3<- glm(cbind(Positives, Total-Positives)~AverageSpringTemp, data = chytrid_binomial, family = "binomial")
summary(M3)

```
Personal exercise 1: Check the number of bulbul specimens per species available on AVONET
```{r}
AVONET <- read.csv("../data/AVONET raw data.csv", stringsAsFactors = FALSE)
str(AVONET)

bulbul_genera <- c("Pycnonotus", "Brachypodius", "Hypsipetes", "Iole",
       "Criniger", "Hemixos", "Alophoixus", "Rubigula",
       "Eurillas", "Setornis", "Tricholestes", "Phyllastrephus",
       "Bleda", "Andropadus", "Stelgidillas", "Calyptocichla", 
       "Neolestes", "Atimastillas", "Alcurus", "Nok", "Spizixos",
       "Ixonotus", "Thescelocichla", "Chlorocichla", "Baeopogon", 
       "Arizelocichla", "Acritillas", "Ixos", "Microtarsus")

AVONET$Species2_eBird <- trimws(as.character(AVONET$Species2_eBird))
pattern <- paste0("^(", paste(bulbul_genera, collapse = "|"), ")(\\b|\\s)")

idx <- !is.na(AVONET$Species2_eBird) & grepl(pattern, AVONET$Species2_eBird, ignore.case = TRUE)
bulbuls <- AVONET[idx, ]
str(bulbuls)

# Only keep species that have at least 4 entries
bulbuls_count <- table(bulbuls$Species2_eBird)
bulbuls <- bulbuls[bulbuls$Species2_eBird %in% names(bulbuls_count[bulbuls_count >= 5]), ]

bulbuls <- read.csv("../data/Bulbul_AVONET_data.csv")
str(bulbuls)
unique(bulbuls$AviList)

bulbuls_var <- bulbuls %>%
  group_by(AviList) %>%
  summarise(
    n_specimens = n(),
    across(
      all_of(trait_cols),
      list(
        Mean = ~ if(sum(!is.na(.)) == 0) NA_real_ else mean(., na.rm = TRUE),
        Max  = ~ if(sum(!is.na(.)) == 0) NA_real_ else max(., na.rm = TRUE),
        Min  = ~ if(sum(!is.na(.)) == 0) NA_real_ else min(., na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

# Save the trait variance data
write.csv(bulbuls_var, "../data/Bulbul_trait_data.csv")

# Ensure Specimen.number is character and trimmed
bulbuls$Specimen.number <- trimws(as.character(bulbuls$Specimen.number))

# Drop rows with missing/empty specimen numbers
bulbuls <- bulbuls[!is.na(bulbuls$Specimen.number) & bulbuls$Specimen.number != "", ]

# Keep only one row per unique specimen number within each species
bulbuls <- bulbuls[!duplicated(paste(bulbuls$Species2_eBird, bulbuls$Specimen.number)), ]
str(bulbuls)

bul_count <- table(bulbuls$Species2_eBird)
mean(bul_count)
hist(bul_count, main = "Frequency of individual specimens sampled per species", xlab = "No. of individuals sampled")

write.csv(bul_count, "../data/Bulbul_specimen_counts_4_or_less.csv")
bul_count <- bul_count[bul_count <= 4]

#Calculate specimen counts again from Bulbul_AVONET_data.csv to check

bul_count <- table(bulbul_avonet$AviList)

setdiff(
  unique(avilist_bulbuls$Scientific_name),
  unique(bulbul_avonet$Species1_BirdLife)
)

setdiff(
  unique(bulbul_avonet$Species1_BirdLife),
  unique(avilist_bulbuls$Scientific_name)
)
```

