Personal exercise 1: Check the number of bulbul specimens per species available on AVONET
```{r}
AVONET <- read.csv("../data/AVONET raw data.csv", stringsAsFactors = FALSE)
str(AVONET)

bulbul_genera <- c("Pycnonotus", "Brachypodius", "Hypsipetes", "Iole",
       "Criniger", "Hemixos", "Alophoixus", "Rubigula",
       "Eurillas", "Setornis", "Tricholestes", "Phyllastrephus",
       "Bleda", "Andropadus", "Stelgidillas", "Calyptocichla", 
       "Neolestes", "Atimastillas", "Alcurus", "Nok", "Spizixos",
       "Ixonotus", "Thescelocichla", "Chlorocichla", "Baeopogon", 
       "Arizelocichla", "Acritillas", "Ixos", "Microtarsus")

AVONET$Species2_eBird <- trimws(as.character(AVONET$Species2_eBird))
pattern <- paste0("^(", paste(bulbul_genera, collapse = "|"), ")(\\b|\\s)")

idx <- !is.na(AVONET$Species2_eBird) & grepl(pattern, AVONET$Species2_eBird, ignore.case = TRUE)
bulbuls_raw <- AVONET[idx, ]
str(bulbuls_raw)

# Only keep species that have at least 4 entries
bulbuls_count <- table(bulbuls_raw$Species2_eBird)
mean(bulbuls_count)
write.csv(bulbuls_count, "../data/Bulbul_RAW_specimen_counts.csv")

bulbuls <- bulbuls_raw[bulbuls_raw$Species2_eBird %in% names(bulbuls_count[bulbuls_count >= 5]), ]

bulbuls <- read.csv("../data/Bulbul_AVONET_data.csv")
str(bulbuls)
unique(bulbuls$AviList)

bulbuls_var <- bulbuls %>%
  group_by(AviList) %>%
  summarise(
    n_specimens = n(),
    across(
      all_of(trait_cols),
      list(
        Mean = ~ if(sum(!is.na(.)) == 0) NA_real_ else mean(., na.rm = TRUE),
        Max  = ~ if(sum(!is.na(.)) == 0) NA_real_ else max(., na.rm = TRUE),
        Min  = ~ if(sum(!is.na(.)) == 0) NA_real_ else min(., na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

# Save the trait variance data
write.csv(bulbuls_var, "../data/Bulbul_trait_data.csv")

# Ensure Specimen.number is character and trimmed
bulbuls$Specimen.number <- trimws(as.character(bulbuls$Specimen.number))

# Drop rows with missing/empty specimen numbers
bulbuls <- bulbuls[!is.na(bulbuls$Specimen.number) & bulbuls$Specimen.number != "", ]

# Keep only one row per unique specimen number within each species
bulbuls <- bulbuls[!duplicated(paste(bulbuls$Species2_eBird, bulbuls$Specimen.number)), ]
str(bulbuls)

bul_count <- table(bulbuls$Species2_eBird)
mean(bul_count)
hist(bul_count, main = "Frequency of individual specimens sampled per species", xlab = "No. of individuals sampled")

write.csv(bul_count, "../data/Bulbul_specimen_counts_4_or_less.csv")
bul_count <- bul_count[bul_count <= 4]

#Calculate specimen counts again from Bulbul_AVONET_data.csv to check

bul_count <- table(bulbuls$AviList)
bul_count <- bul_count[bul_count <= 9]
write.csv(bul_count, "../data/Bulbul_specimen_counts_less_than_10.csv")

setdiff(
  unique(avilist_bulbuls$Scientific_name),
  unique(bulbul_avonet$Species1_BirdLife)
)

setdiff(
  unique(bulbul_avonet$Species1_BirdLife),
  unique(avilist_bulbuls$Scientific_name)
)
```
Looking at holdings of bulbuls in the KU collection
```{r}
ku_holdings <- read.csv("../data/Holdings/ku_pycnonotidae.csv")
str(ku_holdings)

#Combine columns 5 and 6 to get full scientific name and add a space in between
ku_holdings$Scientific_name <- paste(ku_holdings$Genus...Full.Name, ku_holdings$Species...Full.Name, sep = " ")
str(ku_holdings)

ku_bulcount <- table(ku_holdings$Scientific_name)

bulcount <- read.csv("../data/Bulbul_specimen_counts.csv")

#Add the specimen counts from the bulcount data frame to a new column in the ku_bulcount data frame, matching by scientific name
ku_bulcount <- as.data.frame(ku_bulcount)
ku_bulcount <- ku_bulcount |>
  left_join(bulcount, by = c("Var1" = "Var1"))

write.csv(ku_bulcount, "../data/Holdings/KU_priority_list.csv")

#Next holding: WFVZ
wfvz_holdings <- read.xlsx("../data/Holdings/WFVZ Pycnonotidae Study Skins.xlsx")
str(wfvz_holdings)
#If column NEWGENUS is not empty, combine it with column NEWSPECIES to get the full scientific name. If NEWGENUS is empty, use the GENUS and SPECIES columns instead
wfvz_holdings$Scientific_name <- ifelse(
  !is.na(wfvz_holdings$NEWGENUS) & !is.na(wfvz_holdings$NEWSPECIES),
  paste(wfvz_holdings$NEWGENUS, wfvz_holdings$NEWSPECIES, sep = " "),
  ifelse(
    !is.na(wfvz_holdings$NEWGENUS) & is.na(wfvz_holdings$NEWSPECIES),
    paste(wfvz_holdings$NEWGENUS, wfvz_holdings$SPECIES, sep = " "),
    ifelse(
      is.na(wfvz_holdings$NEWGENUS) & !is.na(wfvz_holdings$NEWSPECIES),
      paste(wfvz_holdings$GENUS, wfvz_holdings$NEWSPECIES, sep = " "),
      paste(wfvz_holdings$GENUS, wfvz_holdings$SPECIES, sep = " ")
    )
  )
)

str(wfvz_holdings)

#Make a count table of the number of specimens per species in the wfvz_holdings data frame
wfvz_bulcount <- table(wfvz_holdings$Scientific_name)

#Check differences between the scientific names in the wfvz_holdings data frame and ku_holdings data frame
setdiff(
  unique(wfvz_holdings$Scientific_name),
  unique(ku_bulcount$Var1)
)

setdiff(
  unique(ku_bulcount$Var1),
  unique(wfvz_holdings$Scientific_name)
)

#Now only keep the species present in both data frames in the wfvz_holdings data frame
wfvz_holdings <- wfvz_holdings[wfvz_holdings$Scientific_name %in% unique(ku_bulcount$Var1), ]
str(wfvz_holdings)

#Priority list for KU
ku_priority_list <- read.csv("../data/Holdings/KU_priority_list.csv")
str(ku_priority_list)

#Now only keep species present in the wfvz_holdings data frame in the ku_priority_list data frame
ku_priority_list <- ku_priority_list[ku_priority_list$Species %in% unique(wfvz_holdings$Scientific_name), ]
str(ku_priority_list)

#Replace the Specimen.count.in.KU.collections column in the ku_priority_list data frame with the count of specimens per species in the wfvz_bulcount data frame, matching by scientific name
ku_priority_list$Specimen.count.in.KU.collections <- wfvz_bulcount[ku_priority_list$Species]
str(ku_priority_list)

#Store the updated priority list as a csv file
write.csv(ku_priority_list, "../data/Holdings/WFVZ_priority_list.csv")

setdiff(
  unique(ku_priority_list$Species),
  unique(wfvz_holdings$Scientific_name)
)

setdiff(
  unique(wfvz_holdings$Scientific_name),
  unique(ku_priority_list$Species)
)
```
Now try similar for AMNH
```{r}
lines <- readLines("../data/Holdings/AMNH_bulbul_holdings.csv")

# Split each line by tab
split_lines <- strsplit(lines, "\t")

# Extract the 14th column (between 14th and 15th tab = field 14)
species <- sapply(split_lines, function(x) {
  if (length(x) >= 14) x[14] else NA
})


# Clean up
species <- species[!is.na(species) & species != ""]

#Remove every 3rd word and add to a new column of dataframe called Subspecies
species_df <- data.frame(Species = species)
species_count <- table(species_df$Species)
write.csv(species_count, "../data/Holdings/Raw_AMNH_bulbul_specimen_counts.csv")
species_df$Subspecies <- sapply(species_df$Species, function(x) {
  words <- strsplit(x, " ")[[1]]
  if (length(words) > 2) {
    subspecies <- words[3]
    species <- paste(words[1], words[2])
    return(subspecies)
  } else {    return(NA)
  }
})


head(species_df)

#Keep only the Species columns rows that match with the AviList column in the bulbuls dataset
species_df <- species_df[species_df$Species %in% unique(bulbuls$AviList), ]
str(species_df)

#Count the number of specimens per species in the species_df data frame
amnh_bulcount <- table(species_df$Species)

#Create a new data frame with the species and specimen counts
amnh_bulcount_df <- data.frame(Species = names(amnh_bulcount),
                               Specimen.count.in.AMNH.collections = as.vector(amnh_bulcount))
str(amnh_bulcount_df)

#Now add a new column for comments
amnh_bulcount_df$Comments <- NA

#Now load the bulbul specimen counts file and add a new column to the amnh_bulcount_df data frame with the specimen counts from the bulbul specimen counts file, matching by species name
bulbul_specimen_counts <- read.csv("../data/Bulbul_specimen_counts.csv")
str(bulbul_specimen_counts)
amnh_bulcount_df$Specimen.count.in.AVONET <- bulbul_specimen_counts$Freq[match(amnh_bulcount_df$Species, bulbul_specimen_counts$Var1)]
str(amnh_bulcount_df)

#Now sort from lowest to highest specimen count in AVONET
amnh_bulcount_df <- amnh_bulcount_df[order(amnh_bulcount_df$Specimen.count.in.AVONET), ]
str(amnh_bulcount_df)
write.csv(amnh_bulcount_df, "../data/Holdings/AMNH_priority_list.csv", row.names = FALSE)
```
Now try specimen counts with the integrated dataset
```{r}
bulbuls <- read.csv("../data/Bulbul_AVONET_data.csv")
bulbuls_integrated <- read.csv("../data/Bulbul_AVONET_data_integrated.csv")
bulbul_specimen_counts <- table(bulbuls_integrated$AviList)
mean(bulbul_specimen_counts)
write.csv(bulbul_specimen_counts, "../data/Bulbul_specimen_counts_integrated.csv")
bulbul_specimen_counts <- bulbul_specimen_counts[bulbul_specimen_counts < 20]
write.csv(bulbul_specimen_counts, "../data/Bulbul_specimen_counts_integrated_less_than_20.csv")
unique(bulbuls_integrated$AviList)

#Find AviList species in the bulbuls dataset that are not present in the integrated dataset and vice versa
setdiff(
  unique(bulbuls$AviList),
  unique(rom$AviList)
)

setdiff(
  unique(rom$AviList),
  unique(bulbuls$AviList)
)
```
Now for the Harvard collections, need to make a priority list
```{r}
library(stringr)

lines <- readLines("../data/Holdings/Harvard_bulbul_holdings.csv", warn = FALSE)

# Split by tab
split_lines <- strsplit(lines, "\t")

# Extract species (column 10)
species <- sapply(split_lines, function(x) {
  if (length(x) >= 10) x[10] else NA
})

# Extract subspecies (column 11)
subspecies <- sapply(split_lines, function(x) {
  if (length(x) >= 11) x[11] else NA
})

# Remove header row
species <- species[-1]
subspecies <- subspecies[-1]

# Clean species (remove quotes + ensure only Genus species)
species_clean <- sapply(species, function(x) {
  x <- gsub('"', '', x)
  words <- strsplit(x, " ")[[1]]
  if (length(words) >= 2) {
    paste(words[1], words[2])
  } else {
    NA
  }
})

# Clean subspecies
subspecies_clean <- gsub('"', '', subspecies)
subspecies_clean[subspecies_clean == ""] <- NA

# Combine into dataframe
harvard_df <- data.frame(
  Species = species_clean,
  Subspecies = subspecies_clean
)

# Remove empty rows
harvard_df <- harvard_df[!is.na(harvard_df$Species), ]

head(harvard_df)

harvard_df$Subspecies[is.na(harvard_df$Subspecies)] <- "none"
#Count specimens per species and subspecies
harvard_count <- table(harvard_df$Species, harvard_df$Subspecies)
harvard_count_df <- as.data.frame(harvard_count)
#Remove all the zeros from the count table
harvard_count_df <- harvard_count_df[harvard_count_df$Freq > 0, ]
colnames(harvard_count_df) <- c("Species", "Subspecies", "Specimen.count.in.Harvard.collections")
str(harvard_count_df)
#Now add a new column for comments
harvard_count_df$Comments <- NA
write.csv(harvard_count_df, "../data/Holdings/Harvard_priority_list.csv", row.names = FALSE)
#Arrange by species name
harvard_count_df <- harvard_count_df[order(harvard_count_df$Species), ]
str(harvard_count_df)
#Attach AVONET specimen counts to the harvard_count_df data frame, matching by species name
bulbul_specimen_counts <- read.csv("../data/Bulbul_specimen_counts_integrated.csv")
harvard_count_df$Specimen.count.in.AVONET <- bulbul_specimen_counts$Freq[match(harvard_count_df$Species, bulbul_specimen_counts$Var1)]
str(harvard_count_df)
write.csv(harvard_count_df, "../data/Holdings/Harvard_priority_list.csv", row.names = FALSE)


#Load again
harvard_priority <- read.csv("../data/Holdings/Harvard_priority_list.csv")
str(harvard_priority)

#Arrange by lowest to highest AVONET specimen count
harvard_priority <- harvard_priority[order(harvard_priority$Specimen.count.in.AVONET), ]
str(harvard_priority)
write.csv(harvard_priority, "../data/Holdings/Harvard_priority_list.csv", row.names = FALSE)
```
Need a specimen counts file that instead counts specimens per species required to reach a target of 30 specimens per species in the integrated dataset, excluding species that already have 30 or more specimens
```{r}
bulbul_specimen_counts <- read.csv("../data/Bulbul_specimen_counts_integrated.csv")
str(bulbul_specimen_counts)
bulbul_specimen_counts <- bulbul_specimen_counts[bulbul_specimen_counts$Freq < 30, ]
bulbul_specimen_counts$Specimens.required.to.reach.30 <- 30 - bulbul_specimen_counts$Freq
str(bulbul_specimen_counts)
#Now to remove Freq column and only keep the Species and Specimens.required.to.reach.30 columns while also ordering by highest to lowest number of specimens required
bulbul_specimen_counts <- bulbul_specimen_counts[order(-bulbul_specimen_counts$Specimens.required.to.reach.30), ]
bulbul_specimen_counts <- bulbul_specimen_counts[, c("Var1", "Specimens.required.to.reach.30")]
colnames(bulbul_specimen_counts) <- c("Species", "Specimens.required.to.reach.30")
str(bulbul_specimen_counts)
write.csv(bulbul_specimen_counts, "../data/Bulbul_specimen_counts_to_reach_30.csv", row.names = FALSE)
```
Combining csv files for the Naturalis Leiden holdings
```{r}
# -----------------------------
# 1. Set path to your Downloads folder
# -----------------------------
base_path <- "/Users/brinda/Downloads"

# -----------------------------
# 2. List all part files
# -----------------------------
files <- list.files(
  path = base_path,
  pattern = "bulbul_occurrences_clean_part[0-9]+\\.csv",
  full.names = TRUE
)

# Optional: ensure correct order (part1 → part10)
files <- files[order(as.numeric(gsub(".*part([0-9]+)\\.csv", "\\1", files)))]

# -----------------------------
# 3. Read and combine
# -----------------------------
library(dplyr)

combined_df <- files %>%
  lapply(read.csv) %>%
  bind_rows()

# -----------------------------
# 4. Write final merged file
# -----------------------------
output_path <- file.path(base_path, "Naturalis_bulbul_holdings.csv")

write.csv(combined_df, output_path, row.names = FALSE)

# -----------------------------
# 5. Quick check
# -----------------------------
cat("Merged file saved to:\n", output_path, "\n")
cat("Total rows:", nrow(combined_df), "\n")

#Load it again
naturalis_holdings <- read.csv("../data/Holdings/Naturalis_bulbul_holdings.csv")
str(naturalis_holdings)

#Now only keep the Scientifi_Name and Occurrences column and rename them to Species and Specimen.count.in.Naturalis.collections respectively
naturalis_holdings <- naturalis_holdings %>%
  select(Scientific_Name, Occurrences) %>%
  rename(Species = Scientific_Name, Specimen.count.in.Naturalis.collections = Occurrences)
str(naturalis_holdings)

#Now add new columns for Requested.samples and Comments and fill them with NA
naturalis_holdings$Requested.samples <- NA
naturalis_holdings$Comments <- NA
str(naturalis_holdings)

#Now add a Specimen.count.in.AVONET column by matching the Species column with the Var1 column in the bulbul_specimen_counts data frame and taking the Freq column from the bulbul_specimen_counts data frame
bulbul_specimen_counts <- read.csv("../data/Bulbul_specimen_counts_integrated.csv", stringsAsFactors = FALSE)
str(bulbul_specimen_counts)

# initialize column as NA (preserve NA where no match)
naturalis_holdings$Specimen.count.in.AVONET <- NA_integer_

# match and assign; unmatched remain NA
midx <- match(naturalis_holdings$Species, bulbul_specimen_counts$Var1)
naturalis_holdings$Specimen.count.in.AVONET <- bulbul_specimen_counts$Freq[midx]

str(naturalis_holdings)

#Make specimen count in AVONET appear before Requested.samples and Comments columns
naturalis_holdings <- naturalis_holdings %>%
  select(Species, Specimen.count.in.Naturalis.collections, Specimen.count.in.AVONET, Requested.samples, Comments)
str(naturalis_holdings)

write.csv(naturalis_holdings, "../data/Holdings/Naturalis_priority_list.csv", row.names = FALSE)

#Now load the naturalis priority list again and arrange by lowest to highest specimen count in AVONET
naturalis_priority <- read.csv("../data/Holdings/Naturalis_priority_list.csv")
str(naturalis_priority)
naturalis_priority <- naturalis_priority[order(naturalis_priority$Specimen.count.in.AVONET), ]
str(naturalis_priority)

#Before writing csv, need to fill in the requested samples column with the number of specimens required to reach 30 specimens in AVONET, which is calculated by taking the Specimen.count.in.AVONET column and subtracting it from 30, but only for species that have less than 30 specimens in AVONET. For species that already have 30 or more specimens in AVONET, the Requested.samples column should be filled with 0.
naturalis_priority$Requested.samples <- ifelse(
  naturalis_priority$Specimen.count.in.AVONET < 30,
  pmin(
    30 - naturalis_priority$Specimen.count.in.AVONET,
    naturalis_priority$Specimen.count.in.Naturalis.collections
  ),
  0
)
str(naturalis_priority)

write.csv(naturalis_priority, "../data/Holdings/Naturalis_priority_list.csv", row.names = FALSE)
```
Now to integrate the NJ Collar datasets and check compatibility of the data
```{r}
njcollar1 <- read.xlsx("../data/Collaborator data/NJ_ bulbuls/Alophoixus ochraceus morphometrics.xlsx")
str(njcollar1)

data_path <- "../data/Collaborator data/NJ_ bulbuls/"

files <- list.files(data_path, pattern = "\\.xlsx?$", full.names = TRUE)

library(readxl)
data_list <- lapply(files, function(f) {
  df <- read.xlsx(f)
  df$Source_file <- basename(f)
  # Ensure Comments, Collection, and Date.collected are always character columns
  for (col in c("Comments", "Collection", "Date.collected", "Specimen.number", "Subspecies", "Locality")) {
    if (!(col %in% names(df))) {
      df[[col]] <- NA_character_
    } else {
      df[[col]] <- as.character(df[[col]])
    }
  }
  df
})

# Combine all data frames into one
njcollar_combined <- bind_rows(data_list)
str(njcollar_combined)
write.csv(njcollar_combined, "../data/Collaborator data/NJ_ bulbuls/NJ_bulbul_morphometrics_combined.csv", row.names = FALSE)
```
Now time to check compatibility and integrate
```{r}
njcollar <- read.csv("../data/Collaborator data/NJ_ bulbuls/NJ_bulbul_morphometrics_combined.csv", stringsAsFactors = FALSE)
str(njcollar)
bulbuls <- read.csv("../data/Bulbul_AVONET_data_integrated.csv", stringsAsFactors = FALSE)
str(bulbuls)

njcollar <- njcollar |>
  rename(
    AviList = AviList,
    Beak.Length_Culmen = Bill_TotalCulmen,
    Beak.Width = Beak.Width,
    Tarsus.Length = Tarsus.Length,
    Wing.Length = Wing.Chord,
    Tail.Length = Tail.Length,
    Specimen.number = Specimen.number,
    Locality = Locality,
    Sex = Sex,
    Subspecies = Subspecies,
    Comments = Comments,
    Measurer = Measurer,
    Source = Collection,
    Date.collected = Date.collected
  )

traits <- c(
"Beak.Length_Culmen",
"Beak.Width",
"Tarsus.Length",
"Wing.Length",
"Tail.Length"
)

#Merging step with AVONET data - do NOT remove any columns from the AVONET data frame, just add new rows for UMich specimens, and leave NA where information/columns are missing in UMich but present in AVONET
bulbuls <- read.csv("../data/Bulbul_AVONET_data_integrated.csv")
bulbuls2 <- bind_rows(
  bulbuls,
  njcollar %>% select(any_of(names(bulbuls)))
) %>%
  select(all_of(names(bulbuls)))
str(bulbuls2)

write.csv(bulbuls2, "../data/Bulbul_AVONET_data_integrated.csv", row.names = FALSE)

#AVONET means for each trait, by species
avonet_means <- bulbuls |>
  group_by(AviList) |>
  summarise(
    across(all_of(traits), ~mean(.x, na.rm=TRUE)),
    .groups="drop"
  )
avonet_points <- merge(
  bulbuls,
  avonet_means,
  by="AviList",
  suffixes=c("_specimen","_mean")
)

#Museum means for each trait, by species
njcollar_means <- njcollar |>
  group_by(AviList) |>
  summarise(
    across(all_of(traits), ~mean(.x, na.rm=TRUE)),
    .groups="drop"
  )
njcollar_points <- merge(
  njcollar,
  avonet_means,
  by="AviList",
  suffixes=c("_njcollar","_mean")
)

njcollar_species <- njcollar_points %>%
  filter(!is.na(Wing.Length_njcollar) | !is.na(Tail.Length_njcollar)) %>%
  pull(AviList) %>%
  unique()

avonet_points_filtered <- avonet_points %>%
    dplyr::filter(AviList %in% njcollar_species)

avonet_long <- avonet_points_filtered %>%
  pivot_longer(
    cols = ends_with("_specimen"),
    names_to = "Trait",
    names_pattern = "(.*)_specimen",
    values_to = "specimen"
  ) %>%
  pivot_longer(
    cols = ends_with("_mean"),
    names_to = "Trait_mean",
    names_pattern = "(.*)_mean",
    values_to = "mean"
  ) %>%
  filter(Trait == Trait_mean) %>%
  select(AviList, Trait, specimen, mean)

njcollar_long <- njcollar_points %>%
  pivot_longer(
    cols = ends_with("_njcollar"),
    names_to = "Trait",
    names_pattern = "(.*)_njcollar",
    values_to = "njcollar"
  ) %>%
  pivot_longer(
    cols = ends_with("_mean"),
    names_to = "Trait_mean",
    names_pattern = "(.*)_mean",
    values_to = "mean"
  ) %>%
  filter(Trait == Trait_mean) %>%
  select(AviList, Trait, njcollar, mean)

ggplot() +

  geom_point(
    data = avonet_long,
    aes(mean, specimen, colour = AviList),
    alpha = 0.4
  ) +

  geom_point(
    data = njcollar_long,
    aes(mean, njcollar),
    colour = "#D55E00",
    shape = 17,
    alpha = 0.4,
    size = 1
  ) +

  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dashed"
  ) +

  facet_wrap(~Trait, scales = "free") +

  theme_classic() +
  theme(legend.position = "bottom") +

  labs(
    x = "AVONET species mean",
    y = "Specimen measurement"
  )

#Remove any co-ordinates > 3SD from the 1:1 line for each trait, where SD is calculated from the AVONET specimen measurements for that trait, across all species
avonet_sd <- bulbuls %>%
  group_by(AviList) %>%
  summarise(
    across(all_of(traits), ~sd(.x, na.rm=TRUE)),
    .groups="drop"
  )

avonet_long <- avonet_long %>%
  left_join(avonet_sd, by = "AviList", suffix = c("", "_sd")) %>%
  mutate(
    distance_from_line = abs(specimen - mean),
    outlier = case_when(
      Trait == "Beak.Length_Culmen" & distance_from_line > 3 * Beak.Length_Culmen_sd ~ TRUE,
      Trait == "Beak.Width" & distance_from_line > 3 * Beak.Width_sd ~ TRUE,
      Trait == "Tarsus.Length" & distance_from_line > 3 * Tarsus.Length_sd ~ TRUE,
      Trait == "Wing.Length" & distance_from_line > 3 * Wing.Length_sd ~ TRUE,
      Trait == "Tail.Length" & distance_from_line > 3 * Tail.Length_sd ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  filter(!outlier)

avonet_long <- avonet_long %>%
  select(AviList, Trait, specimen, mean)

#Re-plot after removing outliers
ggplot() +
  geom_point(
    data = avonet_long,
    aes(mean, specimen, colour = AviList),
    alpha = 0.4
  ) +

  geom_point(
    data = njcollar_long,
    aes(mean, njcollar),
    colour = "#D55E00",
    shape = 17,
    alpha = 0.4,
    size = 1
  ) +

  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dashed"
  ) +

  facet_wrap(~Trait, scales = "free") +

  theme_classic() +
  theme(legend.position = "bottom") +

  labs(
    x = "AVONET species mean",
    y = "Specimen measurement"
  )

ggsave("../results/Collaborator data comparisons/njcollar_comparison.pdf", width = 12, height = 8)
```
I need to now make a very good looking phylogeny figure with initial AVONET specimen counts for bulbuls and current bulbul specimen counts after integrating data from all museums
```{r}
#Specimen counts first (bulbuls_raw already loaded)

bulbuls <- read.csv(
  "../data/Bulbul_AVONET_data_integrated.csv",
  stringsAsFactors = FALSE)

traits <- c(
  "Beak.Length_Culmen",
  "Beak.Width",
  "Beak.Depth",
  "Tarsus.Length",
  "Wing.Length",
  "Tail.Length"
)

bulbuls$n_traits_measured <- rowSums(
  !is.na(bulbuls[, traits])
)

bulbuls_filtered <- bulbuls %>%
  filter(n_traits_measured >= 3)
str(bulbuls_filtered)

bulbul_specimen_counts <- bulbuls_filtered %>%
  count(AviList)
write.csv(bulbul_specimen_counts, "../data/Bulbul_specimen_counts_integrated_3traits.csv", row.names = FALSE)
str(bulbul_specimen_counts)

bulbul_specimen_counts <- read.csv("../data/Bulbul_specimen_counts_integrated_3traits.csv", stringsAsFactors = FALSE)

raw_specimen_counts <- read.csv("../data/Bulbul_RAW_specimen_counts.csv", stringsAsFactors = FALSE)
str(raw_specimen_counts)

write.csv(bulbuls_raw, "../data/Bulbul_RAW_AVONET_data.csv", row.names = FALSE)

#Phylogeny time
library(ape)
phylogeny <- read.tree("../data/Phylogeny/mcc_dated_clements_McTavish_2025.nex")
str(phylogeny)

phylogeny <- phylogeny[[6]]

#Now remove the underscores from the tip labels in the phylogeny and replace them with spaces
phylogeny$tip.label <- gsub("_", " ", phylogeny$tip.label)

bulbuls <- read.csv("../data/Bulbul_AVONET_data_integrated.csv", stringsAsFactors = FALSE)

#Now only keep the tips from the phylogeny that are present in the AviList column of the bulbuls data frame
phylogeny <- drop.tip(phylogeny, setdiff(phylogeny$tip.label, unique(bulbuls$AviList)))
str(phylogeny)

#Now create a data frame with the tip labels and the specimen counts from the bulbul_specimen_counts data frame, matching by species name
bulbul_specimen_counts <- read.csv("../data/Bulbul_specimen_counts_integrated_3traits.csv", stringsAsFactors = FALSE)
raw_specimen_counts <- read.csv("../data/Bulbul_RAW_specimen_counts.csv", stringsAsFactors = FALSE)

#Find the different species between the bulbul_specimen_counts and raw_specimen_counts data frames
setdiff(
  unique(bulbul_specimen_counts$AviList),
  unique(raw_specimen_counts$Var1)
)

setdiff(
  unique(raw_specimen_counts$Var1),
  unique(bulbul_specimen_counts$AviList)
)

#Now add the rows for the species that are in bulbul_specimen_counts but not in raw_specimen_counts to the raw_specimen_counts data frame, with a Freq of 0
missing_species <- setdiff(
  unique(bulbul_specimen_counts$Var1),
  unique(raw_specimen_counts$Var1)
)

missing_rows <- data.frame(X = NA, Var1 = missing_species, Freq = 0)
raw_specimen_counts <- rbind(raw_specimen_counts, missing_rows)
write.csv(raw_specimen_counts, "../data/Bulbul_RAW_specimen_counts.csv", row.names = FALSE)


phylo_data <- data.frame(
  Species = phylogeny$tip.label,
  Specimen.count.in.AVONET = raw_specimen_counts$Freq[match(phylogeny$tip.label, raw_specimen_counts$Var1)],
  Specimen.count.after.integration = bulbul_specimen_counts$n[match(phylogeny$tip.label, bulbul_specimen_counts$AviList)]
)
phylo_data$Genus <- sub(" .*", "", phylo_data$Species)
str(phylo_data)

#Now to make a nice phylogeny figure with the initial AVONET specimen counts and the current bulbul specimen counts after integrating data from all museums, as a stacked bar plot coming from the tip labels using ggtree
library(ggtree)
library(ggplot2)
library(ggstance)

#PLOTTING

phylo_data_before <- phylo_data %>%
  select(Species, Genus, Count = Specimen.count.in.AVONET)

phylo_data_after <- phylo_data %>%
  select(Species, Genus, Count = Specimen.count.after.integration)

make_tree_plot <- function(data,
                           pwidth = 1.2,
                           tree_linewidth = 0.5,
                           tip_size = 1.2) {
  
  bar_scale <- 10
  
  data_tree <- data
  
  data_bar <- data %>%
    mutate(
      Count_bar = Count * bar_scale
    )
  
  p <- ggtree(
    phylogeny,
    layout = "fan",
    open.angle = 5,
    size = tree_linewidth
  ) %<+% data_tree
  
  p <- p + geom_tippoint(
    aes(color = Genus),
    size = tip_size
  )
  
  p <- p + geom_fruit(
    data = data_bar,
    geom = geom_bar,
    mapping = aes(
      y = Species,
      x = Count_bar
    ),
    stat = "identity",
    orientation = "y",
    offset = 0.05,
    width = 0.8,
    pwidth = pwidth,
    fill = "steelblue"
  )
  
  p + 
    theme_void() +
    theme(
      legend.position = "right"
    )
}

p_before <- make_tree_plot(
  phylo_data_before,
  pwidth = 0.27,
  tree_linewidth = 0.75,
  tip_size = 2.1
)

p_after <- make_tree_plot(
  phylo_data_after,
  pwidth = 1.2,
  tree_linewidth = 0.5,
  tip_size = 1.2
)

library(cowplot)

legend <- get_legend(
  p_before +
    theme(legend.position = "right") +
    guides(color = guide_legend(title = "Genus"))
)

p_before <- p_before +
  theme(legend.position = "none")

p_after <- p_after +
  theme(legend.position = "none")

final_plot <- plot_grid(
  p_before,
  p_after,
  legend,
  ncol = 3,
  rel_widths = c(1, 1, 0.5)
)

ggsave(
  "../results/Phylogeny/phylogeny_specimen_counts_comparison.pdf",
  final_plot,
  width = 16,
  height = 12
)
```
Now let's try some density plots to compare the distribution of specimen measurements in AVONET 1.0 vs the integrated dataset
```{r}
improvement_df <- phylo_data %>%
  mutate(
    diff = Specimen.count.after.integration -
      Specimen.count.in.AVONET
  )

bulbuls <- read.csv(
  "../data/Bulbul_AVONET_data_integrated.csv",
  stringsAsFactors = FALSE
)

bulbuls$Dataset <- "After"

raw_bulbuls <- read.csv(
  "../data/Bulbul_RAW_AVONET_data.csv",
  stringsAsFactors = FALSE
)

raw_bulbuls$Dataset <- "Before"

raw_bulbuls <- raw_bulbuls %>%
  rename(AviList = Species1_BirdLife)

combined_data <- bind_rows(raw_bulbuls, bulbuls)

plot_data <- combined_data %>%
  filter(!is.na(Wing.Length))

plot_summary <- plot_data %>%
  group_by(AviList, Dataset) %>%
  summarise(
    mean_trait = mean(Wing.Length, na.rm = TRUE),
    sample_size = n(),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(
    names_from = Dataset,
    values_from = c(mean_trait, sample_size)
  ) %>%
  mutate(
    mean_difference = abs(mean_trait_After - mean_trait_Before)
  )

species_selection <- improvement_df %>%
  select(
    Species,
    improvement = diff
  ) %>%
  left_join(
    plot_summary,
    by = c("Species" = "AviList")
  ) %>%
  filter(
    !is.na(mean_difference),
    mean_difference >= 3,
    sample_size_Before >= 3,
    sample_size_After >= 10
  ) %>%
  arrange(
    desc(improvement),
    desc(mean_difference)
  )

top_species <- head(species_selection$Species, 6)

plot_data_filtered <- plot_data %>%
  filter(AviList %in% top_species)

plot_data_filtered$AviList <- factor(
  plot_data_filtered$AviList,
  levels = top_species
)

mean_lines <- plot_data_filtered %>%
  group_by(AviList, Dataset) %>%
  summarise(
    mean_trait = mean(Wing.Length, na.rm = TRUE),
    .groups = "drop"
  )

binwidth_value <- 2

p <- ggplot(
  plot_data_filtered,
  aes(
    x = Wing.Length,
    fill = Dataset
  )
) +
  
  geom_histogram(
    aes(y = after_stat(count)),
    position = "identity",
    alpha = 0.5,
    color = "black",
    linewidth = 0.2,
    binwidth = binwidth_value
  ) +
  
  geom_vline(
    data = mean_lines,
    aes(
      xintercept = mean_trait,
      color = Dataset
    ),
    linewidth = 1,
    linetype = "dashed"
  ) +
  
  facet_wrap(
    ~ AviList,
    scales = "free_y",
    ncol = 3
  ) +
  
  scale_fill_manual(
    values = c(
      "Before" = "red",
      "After" = "steelblue"
    )
  ) +
  
  scale_color_manual(
    values = c(
      "Before" = "darkred",
      "After" = "steelblue4"
    )
  ) +
  
  labs(
    x = "Wing Length (mm)",
    y = "Frequency",
    fill = "Dataset",
    color = "Mean"
  ) +
  
  theme_classic(base_size = 13) +
  
  theme(
    strip.text = element_text(
      face = "italic",
      size = 11
    ),
    legend.position = "top"
  )

p

ggsave(
  "../results/Phylogeny/histograms_top_species.pdf",
  p,
  width = 14,
  height = 9
)

species_selection %>%
  select(
    Species,
    improvement,
    mean_trait_Before,
    mean_trait_After,
    sample_size_Before,
    sample_size_After
  ) %>%
  head(10)
```
Re-trying the phylogeny figure as one plot
```{r}
phylo_data_before <- phylo_data %>%
  select(
    Species,
    Genus,
    Count = Specimen.count.in.AVONET
  )

phylo_data_after <- phylo_data %>%
  select(
    Species,
    Genus,
    Count = Specimen.count.after.integration
  )

global_max <- max(
  c(
    phylo_data_before$Count,
    phylo_data_after$Count
  ),
  na.rm = TRUE
)

make_tree_plot <- function(data) {
  
  data_tree <- data
  
  data_bar <- data %>%
    mutate(
      Count_bar = Count
    )
  
  p <- ggtree(
    phylogeny,
    layout = "fan",
    open.angle = 5,
    size = 0.3
  ) %<+% data_tree
  
  p <- p +
    geom_tippoint(
      aes(color = Genus),
      size = 1.2
    )
  
  p <- p +
    geom_fruit(
      data = data_bar,
      geom = geom_bar,
      mapping = aes(
        y = Species,
        x = Count_bar
      ),
      stat = "identity",
      orientation = "y",
      offset = 0.05,
      width = 0.8,
      pwidth = 2.5,
      fill = "steelblue"
    )
  
  p <- p +
    scale_x_continuous(
      limits = c(0, global_max)
    )
  
  p +
    theme_void() +
    theme(
      legend.position = "right"
    )
}

p_before <- make_tree_plot(phylo_data_before)

p_after <- make_tree_plot(phylo_data_after)

legend <- get_legend(
  p_before +
    theme(legend.position = "right") +
    guides(
      color = guide_legend(title = "Genus")
    )
)

p_before <- p_before +
  theme(legend.position = "none")

p_after <- p_after +
  theme(legend.position = "none")

final_plot <- plot_grid(
  p_before,
  p_after,
  legend,
  ncol = 3,
  rel_widths = c(1, 1, 0.4)
)
```
Last plot needed for meeting: A museum map with all data source lat-longs, then the radius of circle is determined by number of specimens they provided
```{r}

```
