Part 1: Bulbul locality data merging with specimen data
```{r}
library(tidyverse)

str(bulbuls) # uses bulbuls data compiled using bulbul_morphospace.Rmd

localities <- read.csv("../data/Locality_data.csv")
str(localities)

# Merge bulbuls data with locality data
localities_selected <- localities %>%
  select(Specimen.number, Locality.on.label, Locality.edited)
str(localities_selected)
colnames(localities_selected)[2] <- "Locality"

localities_selected <- localities %>%
  filter(!is.na(Specimen.number)) %>%
  distinct(Specimen.number, .keep_all = TRUE)

merged <- bulbuls %>%
  left_join(
    localities_selected,
    by = "Specimen.number",
    suffix = c("", ".new")
  )

merged <- merged %>%
  mutate(
    Locality = ifelse(
      !is.na(Locality.on.label),
      Locality.on.label,
      Locality
    )
  ) %>%
  select(-Locality.on.label)
str(merged)

#Combining Locality.edited column unless NA or empty string
merged <- merged %>%
  mutate(
    Locality = ifelse(
      !is.na(Locality.edited) & Locality.edited != "",
      Locality.edited,
      Locality
    )
  ) %>%
  select(-Locality.edited)
str(merged)

write.csv(merged, "../data/Bulbul_merged_localities.csv")
```

Part 2: Georeferencing bulbul specimen localities
```{r}
localities_unique <- merged %>%
  filter(!is.na(Locality) & Locality != "") %>%
  group_by(Species2_eBird, Locality) %>%
  summarise(
    Specimen.number = paste(Specimen.number, collapse = ", "),
    Country = paste(Country_WRI, collapse = ", "),
    n_specimens = n(),
    .groups = "drop"
  )

str(localities_unique)

write.csv(localities_unique, "../data/Bulbul_unique_localities.csv")
```

Part 3: Put the points on the map
```{r}
unique_localities <- read.csv("../data/Bulbul_unique_localities.csv", stringsAsFactors = FALSE)
str(unique_localities)

library(sf)

points_sf <- st_as_sf(
  unique_localities,
  coords = c("Lon", "Lat"),
  crs = 4326
)

# Plot points to verify
plot(points_sf["n_specimens"])

# Convert n_specimens to factor for discrete color scale
points_sf <- points_sf %>%
  mutate(n_specimens = as.factor(n_specimens))

# Get world borders
world <- ne_countries(scale = "medium", returnclass = "sf")

# Get bounding box of points and add buffer
bbox <- st_bbox(points_sf)
buffer_dist <- 5  # degrees
bbox_buffered <- bbox + c(-buffer_dist, -buffer_dist, buffer_dist, buffer_dist)
bbox_sf <- st_as_sfc(bbox_buffered)

# Crop world to bounding box
world_cropped <- st_crop(world, bbox_buffered)

# Plot using ggplot2 with zoomed extent
ggplot() +
  geom_sf(data = world_cropped, fill = "lightgrey", color = "black") +
  geom_sf(data = points_sf, aes(color = n_specimens), size = 2) +
  coord_sf(xlim = c(bbox_buffered[1], bbox_buffered[3]), 
           ylim = c(bbox_buffered[2], bbox_buffered[4]),
           expand = FALSE) +
  theme_minimal() +
  labs(title = "Bulbul Specimen Localities") +
  theme(legend.position = "top", legend.title = element_text(size = 10)) +
  scale_color_viridis_d(name = "Number of Specimens per species")

ggsave("../results/Bulbul_Specimen_Localities_Map.png", width = 8, height = 6)
```
Add the lat and long columns from the bulbul unique localities data frame to the integrated bulbul data frame, matching by specimen numbers
```{r}
unique_localities <- read.csv("../data/Bulbul_unique_localities.csv", stringsAsFactors = FALSE)
str(unique_localities)
bulbuls <- read.csv("../data/Bulbul_AVONET_data_integrated.csv", stringsAsFactors = FALSE)
str(bulbuls)
#First need to separate the specimen numbers in the unique localities data frame into individual rows
localities_long <- unique_localities %>%
  separate_rows(Specimen.number, sep = ", ") %>%
  rename(Specimen.number = Specimen.number)
#Now we can merge the localities data (but only the lat and long columns) with the bulbuls data
bulbuls_with_localities <- bulbuls %>%
  left_join(localities_long %>% select(Specimen.number, Lat, Lon), by = "Specimen.number")
str(bulbuls_with_localities)
write.csv(bulbuls_with_localities, "../data/Bulbul_AVONET_data_integrated_with_localities.csv", row.names = FALSE)
```
Now use bulbul_unique_localities.csv and append all the missing specimen numbers and localities for bulbuls
```{r}
library(stringr)
library(tidyverse)

bulbuls_localities <- read.csv("../data/Bulbul_final_georeferenced_localities_Batch2complete.csv", stringsAsFactors = FALSE)
bulbuls <- read.csv("../data/Bulbul_AVONET_data_integrated_trimmed.csv", stringsAsFactors = FALSE)

#Check for duplicate specimen numbers in bulbuls_localities
bulbuls_localities %>%
  separate_rows(Specimen.number, sep = ",") %>%
  mutate(Specimen.number = str_trim(Specimen.number)) %>%
  count(Specimen.number) %>%
  filter(n > 1)

#Checking which specimen numbers in bulbuls_localities are present more than once in bulbuls
locality_specimens <- bulbuls_localities %>%
  separate_rows(Specimen.number, sep = ",") %>%
  mutate(
    Specimen.number = str_trim(Specimen.number)
  ) %>%
  pull(Specimen.number)

duplicate_matches <- bulbuls %>%
  filter(
    Specimen.number %in% locality_specimens
  ) %>%
  count(Specimen.number) %>%
  filter(n > 1) %>%
  arrange(desc(n))

duplicate_matches

#Remove all the duplicate rows from bulbuls that are in the duplicate_matches dataframe (but keep the first occurrence of each duplicate)
bulbuls <- bulbuls %>%
  group_by(Specimen.number) %>%
  filter(
    !(Specimen.number %in% duplicate_matches$Specimen.number & row_number() > 1)
  ) %>%
  ungroup()

#Final merger
locality_lookup <- bulbuls_localities %>%
  separate_rows(Specimen.number, sep = ",") %>%
  mutate(
    Specimen.number = str_trim(Specimen.number)
  ) %>% rename(Locality_Corrected = LocalityString) %>%
  select(
    Specimen.number,
    Lat,
    Lon,
    Locality_Corrected
  )

## Update bulbuls with georeferenced values from locality_lookup for matching Specimen.number
# Use left_join and coalesce to update Lat, Lon, Locality_Corrected only where NA
bulbuls <- bulbuls %>%
  left_join(
    locality_lookup,
    by = "Specimen.number",
    suffix = c("", ".lookup")
  ) %>%
  mutate(
    Lat = coalesce(Lat, Lat.lookup),
    Lon = coalesce(Lon, Lon.lookup),
    Locality_Corrected = Locality_Corrected.lookup
  ) %>%
  select(-Lat.lookup, -Lon.lookup, -Locality_Corrected.lookup)
str(bulbuls)
write.csv(bulbuls, "../data/Bulbul_AVONET_data_integrated_trimmed.csv", row.names = FALSE)

##Before I get a new batch, I need to add a column to the bulbuls dataframe that is Locality_Corrected, which is the Locality values from the bulbuls_localities dataframe, matched by Specimen.number
bulbuls <- bulbuls %>%
  left_join(
    locality_lookup %>% rename(Locality_Corrected = LocalityString),
    by = "Specimen.number"
  )

# Create new georeferencing batch
new_localities <- bulbuls %>%
  filter(
    is.na(Lat),
    !is.na(Locality),
    Locality != ""
  ) %>%
  mutate(
    Locality = str_squish(Locality)
  ) %>%
  group_by(
    Locality,
    Country = Country_WRI,
    Source
  ) %>%
  summarise(
    Specimen.number = paste(unique(Specimen.number), collapse = ", "),
    AviList = paste(unique(AviList), collapse = ", "),
    n_specimens = n(),
    .groups = "drop"
  ) %>%
  mutate(
    State = "",
    County = "",
    Lat = NA_real_,
    Lon = NA_real_,
    `Correction Status` = "",
    Precision = "",
    `Error Polygon` = "",
    `Error Polygon (circular polygon)` = "",
    Multiplot = "",
    Code = NA_character_
  ) %>%
  rename(
    LocalityString = Locality
  ) %>%
  select(
    LocalityString,
    Country,
    State,
    County,
    Lat,
    Lon,
    `Correction Status`,
    Precision,
    `Error Polygon`,
    `Error Polygon (circular polygon)`,
    Multiplot,
    AviList,
    Source,
    Specimen.number,
    n_specimens,
    Code
  )

str(new_localities)

write.csv(new_localities, "../data/Bulbul_final_georeferenced_localities_Batch3.csv", row.names = FALSE)
```
Just checking which locality lat long bins would be best for the localiity-based random effect
```{r}
bulbul <- read.csv("../data/Bulbul_plotting_1Jul.csv", stringsAsFactors = FALSE)
str(bulbul)

locality_cluster_summary <- function(
  data,
  resolution
){

  dat_locality <-
    data %>%
    filter(
      !is.na(Lat),
      !is.na(Lon),
      !is.na(AviList)
    ) %>%
    mutate(

      lat_cluster =
        round(
          Lat / resolution
        ) * resolution,

      lon_cluster =
        round(
          Lon / resolution
        ) * resolution,

      locality_cluster =
        interaction(
          AviList,
          lat_cluster,
          lon_cluster,
          drop = TRUE
        )

    )

  cluster_sizes <-
    dat_locality %>%
    count(
      AviList,
      locality_cluster,
      name = "n_specimens"
    )

  species_clusters <-
    cluster_sizes %>%
    group_by(
      AviList
    ) %>%
    summarise(

      n_clusters =
        n(),

      n_singleton_clusters =
        sum(
          n_specimens == 1
        ),

      prop_singleton_clusters =
        mean(
          n_specimens == 1
        ),

      .groups = "drop"

    )

  tibble(

    resolution =
      resolution,

    n_specimens =
      nrow(dat_locality),

    n_species =
      n_distinct(
        dat_locality$AviList
      ),

    n_clusters =
      nrow(cluster_sizes),

    n_singleton_clusters =
      sum(
        cluster_sizes$n_specimens == 1
      ),

    prop_singleton_clusters =
      mean(
        cluster_sizes$n_specimens == 1
      ),

    n_specimens_in_singleton_clusters =
      sum(
        cluster_sizes$n_specimens[
          cluster_sizes$n_specimens == 1
        ]
      ),

    prop_specimens_in_singleton_clusters =
      sum(
        cluster_sizes$n_specimens[
          cluster_sizes$n_specimens == 1
        ]
      ) /
      sum(
        cluster_sizes$n_specimens
      ),

    median_cluster_size =
      median(
        cluster_sizes$n_specimens
      ),

    mean_cluster_size =
      mean(
        cluster_sizes$n_specimens
      ),

    max_cluster_size =
      max(
        cluster_sizes$n_specimens
      ),

    median_clusters_per_species =
      median(
        species_clusters$n_clusters
      ),

    mean_clusters_per_species =
      mean(
        species_clusters$n_clusters
      ),

    n_species_one_cluster =
      sum(
        species_clusters$n_clusters == 1
      ),

    prop_species_one_cluster =
      mean(
        species_clusters$n_clusters == 1
      )

  )

}

cluster_resolutions <-
  c(
    0.1,
    0.25,
    0.5,
    1
  )

cluster_comparison <-
  map_dfr(

    cluster_resolutions,

    ~ locality_cluster_summary(
      data = bulbul,
      resolution = .x
    )

  )

cluster_comparison

get_cluster_sizes <- function(
  data,
  resolution
){

  data %>%
    filter(
      !is.na(Lat),
      !is.na(Lon),
      !is.na(AviList)
    ) %>%
    mutate(

      lat_cluster =
        round(
          Lat / resolution
        ) * resolution,

      lon_cluster =
        round(
          Lon / resolution
        ) * resolution,

      locality_cluster =
        interaction(
          AviList,
          lat_cluster,
          lon_cluster,
          drop = TRUE
        )

    ) %>%
    count(
      AviList,
      locality_cluster,
      name = "n_specimens"
    )

}

clusters_01 <-
  get_cluster_sizes(
    bulbul,
    0.1
  )

clusters_025 <-
  get_cluster_sizes(
    bulbul,
    0.25
  )

clusters_05 <-
  get_cluster_sizes(
    bulbul,
    0.5
  )

clusters_1 <-
  get_cluster_sizes(
    bulbul,
    1
  )

summary(
  clusters_01$n_specimens
)

summary(
  clusters_025$n_specimens
)

summary(
  clusters_05$n_specimens
)

summary(
  clusters_1$n_specimens
)

small_cluster_comparison <-
  map_dfr(

    cluster_resolutions,

    function(res){

      x <-
        get_cluster_sizes(
          bulbul,
          res
        )

      tibble(

        resolution =
          res,

        prop_n1 =
          mean(
            x$n_specimens == 1
          ),

        prop_n2 =
          mean(
            x$n_specimens == 2
          ),

        prop_n3 =
          mean(
            x$n_specimens == 3
          ),

        prop_n_le_2 =
          mean(
            x$n_specimens <= 2
          ),

        prop_n_le_5 =
          mean(
            x$n_specimens <= 5
          )

      )

    }

  )

small_cluster_comparison

```