Part 1: Bulbul locality data merging with specimen data
```{r}
library(tidyverse)

str(bulbuls) # uses bulbuls data compiled using bulbul_morphospace.Rmd

localities <- read.csv("../data/Locality_data.csv")
str(localities)

# Merge bulbuls data with locality data
localities_selected <- localities %>%
  select(Specimen.number, Locality.on.label, Locality.edited)
str(localities_selected)
colnames(localities_selected)[2] <- "Locality"

localities_selected <- localities %>%
  filter(!is.na(Specimen.number)) %>%
  distinct(Specimen.number, .keep_all = TRUE)

merged <- bulbuls %>%
  left_join(
    localities_selected,
    by = "Specimen.number",
    suffix = c("", ".new")
  )

merged <- merged %>%
  mutate(
    Locality = ifelse(
      !is.na(Locality.on.label),
      Locality.on.label,
      Locality
    )
  ) %>%
  select(-Locality.on.label)
str(merged)

#Combining Locality.edited column unless NA or empty string
merged <- merged %>%
  mutate(
    Locality = ifelse(
      !is.na(Locality.edited) & Locality.edited != "",
      Locality.edited,
      Locality
    )
  ) %>%
  select(-Locality.edited)
str(merged)

write.csv(merged, "../data/Bulbul_merged_localities.csv")
```

Part 2: Georeferencing bulbul specimen localities
```{r}
localities_unique <- merged %>%
  filter(!is.na(Locality) & Locality != "") %>%
  group_by(Species2_eBird, Locality) %>%
  summarise(
    Specimen.number = paste(Specimen.number, collapse = ", "),
    Country = paste(Country_WRI, collapse = ", "),
    n_specimens = n(),
    .groups = "drop"
  )

str(localities_unique)

write.csv(localities_unique, "../data/Bulbul_unique_localities.csv")
```

Part 3: Put the points on the map
```{r}
unique_localities <- read.csv("../data/Bulbul_unique_localities.csv", stringsAsFactors = FALSE)
str(unique_localities)

library(sf)

points_sf <- st_as_sf(
  unique_localities,
  coords = c("Lon", "Lat"),
  crs = 4326
)

# Plot points to verify
plot(points_sf["n_specimens"])

# Convert n_specimens to factor for discrete color scale
points_sf <- points_sf %>%
  mutate(n_specimens = as.factor(n_specimens))

# Get world borders
world <- ne_countries(scale = "medium", returnclass = "sf")

# Get bounding box of points and add buffer
bbox <- st_bbox(points_sf)
buffer_dist <- 5  # degrees
bbox_buffered <- bbox + c(-buffer_dist, -buffer_dist, buffer_dist, buffer_dist)
bbox_sf <- st_as_sfc(bbox_buffered)

# Crop world to bounding box
world_cropped <- st_crop(world, bbox_buffered)

# Plot using ggplot2 with zoomed extent
ggplot() +
  geom_sf(data = world_cropped, fill = "lightgrey", color = "black") +
  geom_sf(data = points_sf, aes(color = n_specimens), size = 2) +
  coord_sf(xlim = c(bbox_buffered[1], bbox_buffered[3]), 
           ylim = c(bbox_buffered[2], bbox_buffered[4]),
           expand = FALSE) +
  theme_minimal() +
  labs(title = "Bulbul Specimen Localities") +
  theme(legend.position = "top", legend.title = element_text(size = 10)) +
  scale_color_viridis_d(name = "Number of Specimens per species")

ggsave("../results/Bulbul_Specimen_Localities_Map.png", width = 8, height = 6)
```
Add the lat and long columns from the bulbul unique localities data frame to the integrated bulbul data frame, matching by specimen numbers
```{r}
unique_localities <- read.csv("../data/Bulbul_unique_localities.csv", stringsAsFactors = FALSE)
str(unique_localities)
bulbuls <- read.csv("../data/Bulbul_AVONET_data_integrated.csv", stringsAsFactors = FALSE)
str(bulbuls)
#First need to separate the specimen numbers in the unique localities data frame into individual rows
localities_long <- unique_localities %>%
  separate_rows(Specimen.number, sep = ", ") %>%
  rename(Specimen.number = Specimen.number)
#Now we can merge the localities data (but only the lat and long columns) with the bulbuls data
bulbuls_with_localities <- bulbuls %>%
  left_join(localities_long %>% select(Specimen.number, Lat, Lon), by = "Specimen.number")
str(bulbuls_with_localities)
write.csv(bulbuls_with_localities, "../data/Bulbul_AVONET_data_integrated_with_localities.csv", row.names = FALSE)
```
Now use bulbul_unique_localities.csv and append all the missing specimen numbers and localities for bulbuls
```{r}
library(stringr)

bulbuls_unique_localities <- read.csv("../data/Bulbul_unique_localities.csv", stringsAsFactors = FALSE)
bulbuls <- read.csv("../data/Bulbul_AVONET_data_integrated.csv", stringsAsFactors = FALSE)

# Split specimen numbers into individual rows
geo_specimens <- bulbuls_unique_localities %>%
  separate_rows(Specimen.number, sep = ",") %>%
  mutate(Specimen.number = str_trim(Specimen.number))

bulbuls_remaining <- bulbuls %>%
  filter(!is.na(Specimen.number)) %>%
  filter(!(Specimen.number %in% geo_specimens$Specimen.number))

# Clean locality
bulbuls_remaining <- bulbuls_remaining %>%
  mutate(Locality = str_squish(Locality))

# Create lookup (deduplicated)
bulbuls_lookup <- bulbuls %>%
  select(Specimen.number, AviList) %>%
  distinct(Specimen.number, .keep_all = TRUE)

# Join cleanly
bulbuls_remaining <- bulbuls_remaining %>%
  select(-AviList) %>%   # remove empty column if present
  left_join(bulbuls_lookup, by = "Specimen.number")

# Summarise
new_localities <- bulbuls_remaining %>%
  filter(!is.na(Locality) & Locality != "") %>%
  group_by(Locality, Country) %>%
  summarise(
    Specimen.number = paste(unique(Specimen.number), collapse = ", "),
    AviList = paste(unique(AviList), collapse = ", "),
    n_specimens = n(),
    .groups = "drop"
  )

new_localities <- new_localities %>%
  rename(
    LocalityString = Locality,
    Country = Country
  ) %>%
  mutate(
    State = "",
    County = "",
    Lat = NA,
    Lon = NA,
    `Correction Status` = "",
    Precision = "",
    `Error Polygon` = "",
    `Error Polygon (circular polygon)` = "",
    Multiplot = "",
    AviList = AviList,
    Specimen.number = Specimen.number,
    n_specimens = n_specimens
  ) %>%
  select(
    LocalityString, Country, State, County, Lat, Lon,
    `Correction Status`, Precision, `Error Polygon`,
    `Error Polygon (circular polygon)`, Multiplot,
    AviList, Specimen.number, n_specimens
  )

final_dataset <- bind_rows(
  bulbuls_unique_localities,
  new_localities
)

str(final_dataset)

final_dataset$Code <- NA
write.csv(final_dataset, "../data/Bulbul_final_georeferenced_localities.csv", row.names = FALSE)
```